---
title: "Cleaning_matrix"
author: "Leslie Kollar"
date: "1/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.width = 40, fig.height = 20)
```


```{r libraries, message=FALSE}
library(tidyverse)
library(vegan)
library(plotly)
library(corrplot)
library(ggpubr)
library(Hmisc)
library(vegan)
library(reshape2)
library(parallel)
library(dfoptim)
library(lsmeans)
library(psych)
library(DMwR)
theme_set(theme_bw())
```

#Reading in the raw data
```{r}
master <- read_csv("/Users/lesliekollar/Desktop/Moss_matrix/rawdata/LK_master_2.csv")
master <- master[1:643,]
master$PTR <- NULL
```

#Goals of this code
Here, we are cleaning the entire matrix data set in order to use in the manuscipt

```{r}
# Change variable names and fix sample ID names with spaces, or parentheses

trait_raw <- master[,1:20] %>% 
  rename(famid = Fam,
         sampid = `Sample Name`,
         ssex = Sample_Sex) %>% 
  drop_na(famid) %>% 
  mutate(sampid = str_replace_all(sampid, "\\(.*\\)", ""),
         sampid = str_trim(sampid, side = "both"),
         ssex = str_to_lower(as.character(ssex))) %>% # Sex on lower case only
  mutate_at(vars(starts_with("Leaf")), as.numeric) %>% # Make them all numeric 
  group_by(famid) %>% 
  arrange(sampid, .by_group = TRUE) %>% # Order by family number
  ungroup

# Fix a data entry error:
trait_raw$Avg_21_days[which(trait_raw$Avg_21_days == 33)] <- 0.33

# Fix sex data entry error
trait_raw[which(trait_raw$sampid == "P_18_1_6_B"),]$ssex <- "m"
trait_raw <- trait_raw[-which(trait_raw$sampid == "P_6_6_20"),]

head(trait_raw)

```
```{r}
# Removing DIV/0 from the data set and setting to NA

trait_raw$Leaf_Average_Area <-  as.numeric(str_replace(trait_raw$Leaf_Average_Area, "#DIV/0", "NA"))
trait_raw$Leaf_Length_Average <-  as.numeric(str_replace(trait_raw$Leaf_Length_Average, "#DIV/0", "NA"))
trait_raw$Leaf_perimeter_average <-  as.numeric(str_replace(trait_raw$Leaf_perimeter_average, "#DIV/0", "NA"))
```


# Removing May 23, 29, and 30
-These days were clear outliers
-This can be seen when cleaning the individual data sets.
```{r}
trait_raw %>% 
  filter(`Date PTR`!= "30-May") %>%
  mutate_if(is.numeric, list(~ ifelse(. < 0, 0, .))) %>% 
  arrange(sampid, .by_group = TRUE)-> trait_raw_3

trait_raw_3 %>% 
  filter(`Date PTR` != "29-May") %>%
  mutate_if(is.numeric, list(~ ifelse(. < 0, 0, .))) %>% 
  arrange(sampid, .by_group = TRUE) -> trait_raw_2

trait_raw_2 %>% 
  filter(`Date PTR` != "23-May") %>%
  mutate_if(is.numeric, list(~ ifelse(. < 0, 0, .))) %>% 
  arrange(sampid, .by_group = TRUE) -> trait_raw
```

#### Leaf traits
-For each leaf trait we took ~3 leaf measurements per replicate. These measurements were averaged for each replicate to get one value per replicate.
-First, we visually check for outliers or values that may not make sense. 
-There are NAs in the dataset which will be problematic but some will be taken care of when we average them
- NA's not resolved by averaging clones will be imputed using nearest neighbor. These zeros are not informative and are missing data.
```{r}
# Leaf data averaging 
trait_raw[,18:20] %>% 
  gather(key = "trait", value = "value") %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```
- There are no clear outliers or errors, so these look ok to average and we create a new data frame with the leaf data. 
- In cases where data for only one of the clones was available, we kept that value. 
-For cases where data was available for both clones we averaged it.  

#Averaging clones for leaf data and removing NAs
```{r}
trait_raw[,c(2, 18:20),] %>% 
  group_by(sampid) %>% # sampid is the unique identifier for the clones
  summarise_at(vars(starts_with("Leaf")), mean, na.rm = TRUE) %>% 
  mutate_at(vars(starts_with("Leaf")), list(~ as.numeric(ifelse(. == "NaN", "NA", .)))) %>% 
  drop_na(sampid)-> leaf_data

# Some NAs
leaf_data[!complete.cases(leaf_data), ]

#Viewing the data after averaging clones
leaf_data[,2:4] %>% 
  gather(key = "trait", value = "value") %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()

```



#### Growth and development traits
-We do the same thing with these other traits and it is easy to identify three values that seem incorrect. These are in the area, circularity and perimeter rate.

```{r}
# Growth and Development traits -------------------------------------------

trait_raw[,11:17] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```

#Removing outliers for the growth data
- For days until gametophores were produced, some values are 100 because that sample never produced gametophores. Oringinally we thought to put in a very high value for the number of days however, that was not informative as it skewed the data. Here, we will convert all 100's to "0" and then added 21 days. 21 days is reasonable to use because that was the last day we collected data 
-All other outliers were noticable as most likely a misprint and fixed.
```{r}
outliers_area <- which(trait_raw$Avg_Area_Week_3>50)
trait_raw[outliers_area,]

outliers_circ <- which(trait_raw$Avg_Circularity_Week_3>50)
trait_raw[outliers_circ,]

#Converting what we originally input as 100 to 0. 100 was entered for samples that never produced gametophores but it skews the data so we are using 21 which is the max amount of days. 

#100 -> 0
outliers_Gam <- which(trait_raw$Avg_Days_til_Gam>25)
trait_raw[outliers_Gam,]
trait_raw$Avg_Days_til_Gam[outliers_Gam] <- (trait_raw$Avg_Days_til_Gam[outliers_Gam] == 0) 

#Inputting 21 
outliers_zero <- which(trait_raw$Avg_Days_til_Gam<=0)
trait_raw[outliers_zero,]
trait_raw$Avg_Days_til_Gam[outliers_zero] <- (trait_raw$Avg_Days_til_Gam[outliers_zero] + 21) 


```

#Double checking after removing outliers
These values had mistakes in the number of decimal places and included a typo making a value be negative. After adjusting these values, we check again for errors. 
```{r}
#Removing samples with mistake in decimal place
trait_raw$Avg_Area_Week_3[outliers_area] <- trait_raw$Avg_Area_Week_3[outliers_area] / 1000
trait_raw$Avg_Circularity_Week_3[outliers_circ] <- trait_raw$Avg_Circularity_Week_3[outliers_circ] / 1000


# check again 
trait_raw[,11:17] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```


```{r}
#Removing outliers that were negative which was a typo
negs <- which(trait_raw$Avg_Perim_Rate < 0)
trait_raw[negs,]

trait_raw$Avg_Perim_Rate[negs]
trait_raw$Avg_Perim_Rate[negs] <- trait_raw$Avg_Perim_Rate[negs] /-10

# Check visually once more...
trait_raw[,11:17] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```

#Combining final data for growth and development data
```{r}
trait_raw[,c(2, 11:17)] %>% 
  drop_na() %>% 
  distinct()-> gro_dev_data
```

#### Reproduction variable
-We measured male reproductive values by the number of sex buds on 10 different gametophores. We did the same thing for females however, females are more difficult to find and therefore we could see less female sex structures simply because of the difficulty to identify them. Because of this, it is better to use the number of archegonia per sex structure as the reproductive unit. For this measurement you have to identify the sex bud and then count the number of archegonia. Using this should resolve any differenes due to difficulty finding them. However, they are not comparable between the sexes as is..
-So we can scale them and that way we can keep them in a single column and compare males and females. 
-The first thing we do here is take the average for the clones (meaning we group by the sample ID, take the average) which we call `raw_av` and then scale this so that both male and female values are in the same *scale*, and we call this `reprovar`.

```{r}
# Reproduction variable ---------------------------------------------------
trait_raw %>% 
  filter(ssex == "m") %>% 
  select(sampid, ssex, `Avg_Male_Buds/Stem`)%>% 
  drop_na() %>% 
  group_by(sampid) %>% 
  summarise(raw_av = mean(`Avg_Male_Buds/Stem`)) %>% 
  mutate(reprovar = scale(raw_av),
         ssex = "m")  -> male_reprovar

trait_raw %>% 
  select(sampid, ssex, Avg_Arch) %>% 
  filter(ssex == "f") %>% 
  drop_na() %>% 
  group_by(sampid) %>% 
  summarise(raw_av = mean(Avg_Arch)) %>% 
  mutate(reprovar = scale(raw_av),
         ssex = "f") -> fem_reprovar


reprovar <- bind_rows(fem_reprovar, male_reprovar)
head(reprovar)
reprovar %>% 
  ggplot(aes(x = reprovar, fill = ssex)) +
  geom_histogram()
```

### Clean trait data
-We can now join all the smaller datasets and have a master traits data frame. 
-These are not scaled so far and we can evaluate the changes of scaling them vs keeping them as they are.

```{r}
# Join trait data ---------------------------------------------------------
trait_raw %>% 
  select(famid, sampid, ssex) %>% 
  distinct() %>% 
  full_join(., reprovar) %>% 
  full_join(., gro_dev_data) %>%
  full_join(., leaf_data) %>% 
  set_names(c("famid", "sampid", "ssex", "raw_repro", "repro", "area_wk3",
              "perim_wk3", "circ_wk3", "perim_rate", "area_rate",
              "days21", "days_gam", "leaf_length", "leaf_area",
              "leaf_perim")) -> traits

#Saving raw clean trait data
saveRDS(traits, "cleandata/clean_traits_final.RDS")

```

#### Histograms raw trait data
These are the values for all the traits that were measured and they include only one value per individual (clones averaged). 

```{r}
traits %>% 
  select(-c(famid, sampid, ssex, raw_repro)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")
```

#### Histograms transformed trait data
Transforming the data does add anything and therefore we will not transform this data for future analyses.
```{r}
names(traits)
traits %>% 
  mutate_at(c(5:15), list(~ log10(. + 1))) %>% 
  mutate_at(c(5:15), scale) -> scaled_traits

scaled_traits %>% 
  select(-c(famid, sampid, ssex, raw_repro)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#7e9bcc") +
  facet_wrap( ~ trait, scales = "free")
```


#VOC data cleaning
```{r}
# Just keep volatile data and identifiers
# Also, remove the NAs, there's a lot at the end
# Remove '(2)' notation from sample ids


voc_raw <- master[,c(1,2,5,6, 21:110)] %>% 
  rename(famid = Fam,
         sampid = `Sample Name`,
         ssex = Sample_Sex) %>% 
  drop_na() %>% 
  mutate(sampid = str_replace_all(sampid, "\\(.*\\)", ""),
         sampid = str_trim(sampid, side = "both"),
         ssex = str_to_lower(as.character(ssex)))

#Shorten voc names
oldnames <- names(voc_raw)
names(voc_raw)[5:94] <- stringr::str_trunc(oldnames[5:94], width = 6, side = "right", ellipsis = "")
```

-There was an issue with the notation in one of the compounds, so we check that and fix the notation so that R can recognize the values as numbers.
```{r}
#str(voc_raw)
#' m111.0 is a character
#voc_raw$m111.0
#' The notation is different, using a capital E
voc_raw$m111.0 <- as.numeric(str_replace(voc_raw$m111.0, "E", "e"))
```
# Removing May 23, 29, and 30
These days were clear outliers among all other days of collection
```{r}

voc_raw %>% 
  filter(`Date PTR`!= "30-May") %>%
  mutate_if(is.numeric, list(~ ifelse(. < 0, 0, .))) %>% 
  arrange(sampid, .by_group = TRUE)-> vocs

vocs %>% 
  filter(`Date PTR` != "29-May") %>%
  mutate_if(is.numeric, list(~ ifelse(. < 0, 0, .))) %>% 
  arrange(sampid, .by_group = TRUE) -> vocs_2

vocs_2 %>% 
  filter(`Date PTR` != "23-May") %>%
  mutate_if(is.numeric, list(~ ifelse(. < 0, 0, .))) %>% 
  arrange(sampid, .by_group = TRUE) -> vocs_3
```

# 10% threshold 
-Compounds where less than 10% of the samples produced it were removed from the dataset as those were assumed to be noise.
```{r}

vocs_3 %>% 
  select(-famid, -sampid, -ssex, -`Date PTR`) %>% 
  mutate_if(is.numeric, function(x) ifelse(x > 0, 1, 0)) %>% 
  colSums() %>% 
  data.frame() %>% 
  rownames_to_column() %>%
  setNames(., c("voc", "count")) %>% 
  mutate(prcnt = count/nrow(vocs)) %>% 
  filter(prcnt >= 0.1) -> voc_filter

#vocs_3 %>% 
#  select(famid, sampid, ssex, Date.PTR, c(voc_filter$voc)) %>% 
#  mutate(ssex = str_to_lower(as.character(ssex))) -> voc_data

#Taking the max for each VOC
#The distinct function is not working
vocs_3 %>% 
  select(famid, sampid, ssex, c(voc_filter$voc)) %>% 
  group_by(sampid) %>%
  mutate_at(vars(starts_with("m")), max) %>% # This is selecting the maximum value for all the volatiles
  distinct() -> voc_data
```

```{r}
#m99.08 was multiplied by 5*10^6 and not 5*10^-6. fix this. 
trait_raw$Avg_Perim_Rate[negs] <- trait_raw$Avg_Perim_Rate[negs] /-10
voc_data$m99.08 <- voc_data$m99.08* (10^-12)
```

# Exploration with clones in VOC data set
```{r}
long_voc_data <- voc_data %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))

long_voc_data %>%
  ggplot(aes(x = voc, y = conc)) +
  geom_point() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

long_voc_data %>% 
  filter(conc > 0) %>%
  ggplot(aes(x = voc, y = conc)) + 
  geom_boxplot() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10()

```




#Cut off values for voc data
-We see that there are three big outliers. The values are so small and over the machine's detection thresold, so we will make them equal to zero.
-We used a cut off for the mass flux of 10 e-10 because anything below this could be due to variation from subtracting out the background taken from the averaging multiple blank samples (control) used during that specific batch of samples

```{r}
# Making these outliers be zero
voc_data[voc_data < 1e-15 & voc_data>0] <- 0

#Cuttoff at 10e-10 in original data file. This is a very liberal cut off. 
voc_data[voc_data < 1e-10] <- 0

long_voc_data <- voc_data %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))
  
long_voc_data%>%
filter(conc > 0) %>%
  ggplot(aes(x = voc, y = conc)) + 
  geom_boxplot() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10()



```

```{r eval = FALSE}
#Plotting log data
long_voc_data %>% 
  ggplot(aes(x = voc, y = conc)) + 
  geom_boxplot() +
  theme_bw() + 
  lims(y = c(1e-15, 1e-02)) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10()
```


#Log transformation VOC data
We added a very small value of 1e-12 rather than 1 to avoid skewing the data. There are zeros in the VOC data however, those are uniformative because they could not be detected from the PTR or the sample could not produce them. The clones can tell us some information about this however, we can never be sure. The log transformation can not occur on a zero and therefore a value needs to be added. 
```{r, fig.width=2.75,fig.height=4.5}
#Log transform data but add 1e-12 to zeros because you cannot log transform zeros
log_vocs <- voc_data %>% 
  mutate_at(vars(starts_with("m")), list(~ log10(. + 1e-12)))

#Checking for missing data- no missing cases
log_vocs[complete.cases(log_vocs),]
```

### Master dataframe
- This will create the master data frame that includes the lifehistory traits and the VOC data. 
- Also, there is one row per individual, which means the data for clones has either been averaged(for traits) or the max value has been taken (volatiles).
- We imputed the missing data using nearest neighbor with euclidean distance. This was onle done for leaf measurements and area, perimeter and ciruclarity. These are the only places were zeros didnt make sense. For example a zero in the reproductive traits means that the sample was not sex expressing and therefore a zero would make sense.

```{r}

clean_master <- full_join(traits, voc_data) %>% 
  mutate_at(c(5:32), as.numeric)

#Print master file (nothing is transformed because of zeros introduced by joining)
clean_master$Date.PTR <- NULL
clean_master$area_wk3[clean_master$area_wk3== 0] <- NA
clean_master$perim_wk3[clean_master$perim_wk3== 0] <- NA
clean_master$circ_wk3[clean_master$circ_wk3== 0] <- NA
clean_master$leaf_area[clean_master$leaf_area== 0] <- NA
clean_master$leaf_length[clean_master$leaf_length== 0] <- NA
clean_master$leaf_perim[clean_master$leaf_perim== 0] <- NA

clean_master$raw_repro <- NULL
clean_master$perim_rate <- NULL
clean_master$area_rate <- NULL

clean_master$famid <- as.factor(clean_master$famid) 
clean_master$sampid <- as.factor(clean_master$sampid) 
clean_master$ssex <- as.factor(clean_master$ssex) 

#Saving clean matrix before imputing the data
saveRDS(clean_master, "~/Desktop/Moss_matrix/cleandata/clean_master_matrix.RDS") 

#Imputing missing data with nearest neighbor prior to log transforming VOC data or scaling
clean_master <- as.data.frame(clean_master)
clean_master_2 <-knnImputation(clean_master, k=6, scale= T, meth = "weighAvg", distData = NULL)


#Log transform only VOC data adding smallest value
clean_master_log <- clean_master_2 %>% 
  mutate_at(vars(starts_with("m")), list(~ log10(. + 1e-12)))

#Seperating by sex with log data only

clean_master_log%>%
  filter(ssex=="m")-> male_master_log

clean_master_log%>%
  filter(ssex=="f")-> female_master_log


#Seperating by sex for seperate analyses wwith log and scaled data

clean_master_log%>%
  filter(ssex=="m")%>%
  mutate_at(vars(starts_with("m")), scale)-> male_master_log_scale #This only scales VOC

clean_master_log%>%
  filter(ssex=="f")%>%
  mutate_at(vars(starts_with("m")), scale)-> female_master_log_scale #This only scales VOC

#Scaling all data!
clean_master_log_scaled <- clean_master_log %>% 
  mutate_at(vars(starts_with("r")), scale)

##All of this data is imputed. Earlier in the code "clean matrix" was printed and saved. 
#write.csv(clean_master_log_scaled, "~/Desktop/Moss_matrix/cleandata/clean_master_matrix_scaled_log_imputed.csv") 
saveRDS(clean_master_log_scaled, "~/Desktop/Moss_matrix/cleandata/clean_master_matrix_scaled_log_imputed.RDS") #clean master both log scaled and transformed

saveRDS(clean_master_2, "~/Desktop/Moss_matrix/cleandata/clean_master_matrix_imputed.RDS") #clean master not log transformed and not scaled

saveRDS(clean_master_log, "~/Desktop/Moss_matrix/cleandata/clean_master_matrix_log_voc_imputed.RDS")
write.csv(clean_master_log, "~/Desktop/Moss_matrix/cleandata/clean_master_matrix_log_voc_imputed.csv")
#clean master log transformed but not scaled
```
















