---
title: "Traits and VOCs correlations"
author: "Javiera Rudolph"
date: "6/11/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(corrplot)
library(ggpubr)
```


# Traits


```{r}
raw_traits <- readRDS("cleandata/traits_clean.RDS")

# Just scale the data
# Couldn't log transform: issues with zeroes, or negative numbers. It's a mess
raw_traits %>% 
  mutate_at(c(5:15), list(~ scale(.))) -> transformed_traits
```


```{r}
# How does the transformed data for traits look like?

transformed_traits %>% 
  select(-c(famid, sampid, ssex, raw_repro)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")

```


## MANOVA for traits

```{r}
# Change the traits here to see transformed vs raw.
traits <- raw_traits

# This runs the manova
res.man <- manova(cbind(repro, area_wk3, perim_wk3, circ_wk3, perim_rate, area_rate, days21, days_gam, leaf_length, leaf_area, leaf_perim) ~ famid + ssex + famid:ssex, data = traits)

```

MANOVA results, shows sex and family to be signficant, and the interaction is borderline significant.
```{r}
summary(res.man)
```


Check with anova for each trait
```{r}
summary.aov(res.man)
```

For each trait get the residuals and from here get family level means, for all data or by sex.
```{r}

res_dat <- cbind(res.man$model[,-1], res.man$residuals)

fam.means <- res_dat %>%
  group_by(famid) %>% 
  mutate_at(vars(3:13), mean) %>% 
  ungroup() %>% 
  select(-ssex) %>% 
  distinct()

fem.means <- res_dat %>% 
  filter(ssex == "f") %>% 
  group_by(famid) %>% 
  mutate_at(vars(3:13), mean) %>% 
  ungroup() %>% 
  select(-ssex) %>% 
  distinct()


male.means <- res_dat %>% 
  filter(ssex == "m") %>% 
  group_by(famid) %>% 
  mutate_at(vars(3:13), mean) %>% 
  ungroup() %>% 
  select(-ssex) %>% 
  distinct()

```

# Correlation plots

For all the data, just looking at family means
```{r}
# Correlations for all 
all.corr <- fam.means %>% 
  select(-c(famid)) %>% 
  cor()
corrplot(all.corr, method = "circle", type = "upper", tl.col = "black", tl.cex = 0.8, tl.srt = 45)

```

By sex, with females on the left, males on the right.
```{r fig.width=12, fig.height=7}

par(mfrow = c(1, 2))

# Correlations for females
fem.corr <- fem.means %>% 
  select(-c(famid)) %>% 
  cor()
corrplot(fem.corr, method = "circle", type = "upper", tl.col = "black", tl.cex = 0.8, tl.srt = 45)


# Correlations for males
male.corr <- male.means %>% 
  select(-c(famid)) %>% 
  cor()
corrplot(male.corr, method = "circle", type = "upper", tl.col = "black", tl.cex = 0.8, tl.srt = 45)

```


If we compare males (x axis) and females (y axis)
```{r}

# Correlations between males and females

# First make sure you are comparing the same families.
fem.means %>% 
  filter(famid %in% male.means$famid) -> new.fem.means

ssex.corr <- cor(x = male.means[,-1], y = new.fem.means[,-1])
corrplot(ssex.corr, method = "circle", tl.col = "black", tl.cex = 0.8, tl.srt = 45)
```




# VOCs

```{r}
voc_clust <- readRDS("cleandata/clustered_voc_data.RDS")

voc_clust %>% 
  mutate_at(c(4:18), list(~ log10(. + 1e-13))) %>% 
  mutate_at(c(4:18), list(~ scale(.))) -> transformed_vocs
```

```{r}
# How does the transformed data for VOCs look like?
transformed_vocs %>% 
  select(-c(famid, sampid, ssex)) %>% 
  gather(key = "trait", value = "values") %>% 
  ggplot(aes(x = values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")
```


## MANOVA for VOCs clusters

```{r}

vocs <- transformed_vocs

# Run the manova
res.man.vocs <- manova(cbind(clust_01, clust_02, clust_03, clust_04, clust_05,
                        clust_06, clust_07, clust_08, clust_09, clust_10, 
                        clust_11, clust_12, clust_13, clust_14, clust_15) ~ famid + ssex + famid:ssex, data = vocs)
```

Overall results from the MANOVA, sex and family significant

```{r}
summary(res.man.vocs)
```

by cluster
```{r}
summary.aov(res.man.vocs)
```


```{r}
res_dat.vocs <- cbind(res.man.vocs$model[,-1], res.man.vocs$residuals)

fam.means.vocs <- res_dat.vocs %>%
  group_by(famid) %>% 
  mutate_at(vars(3:17), mean) %>% 
  ungroup() %>% 
  select(-ssex) %>% 
  distinct()

fem.means.vocs <- res_dat.vocs %>% 
  filter(ssex == "f") %>% 
  group_by(famid) %>% 
  mutate_at(vars(3:17), mean) %>% 
  ungroup() %>% 
  select(-ssex) %>% 
  distinct()


male.means.vocs <- res_dat.vocs %>% 
  filter(ssex == "m") %>% 
  group_by(famid) %>% 
  mutate_at(vars(3:17), mean) %>% 
  ungroup() %>% 
  select(-ssex) %>% 
  distinct()

```

### Correlations

For all
```{r}

# Correlations for all 
all.corr.vocs <- fam.means.vocs %>% 
  select(-c(famid)) %>% 
  cor()
corrplot(all.corr.vocs, method = "circle", type = "upper", tl.col = "black", tl.cex = 0.8, tl.srt = 45)

```


Females on the left, males on the right
```{r fig.width=12, fig.height=7}

par(mfrow = c(1, 2))

# Correlations for females
fem.corr.vocs <- fem.means.vocs %>% 
  select(-c(famid)) %>% 
  cor()
corrplot(fem.corr.vocs, method = "circle", type = "upper", tl.col = "black", tl.cex = 0.8, tl.srt = 45)


# Correlations for males
male.corr.vocs <- male.means.vocs %>% 
  select(-c(famid)) %>% 
  cor()
corrplot(male.corr.vocs, method = "circle", type = "upper", tl.col = "black", tl.cex = 0.8, tl.srt = 45)

```

Males on the x axis, females on the y
```{r}

# Correlations between males and females

fem.means.vocs %>% 
  filter(famid %in% male.means.vocs$famid) -> new.fem.means.vocs

ssex.corr.vocs <- cor(x = male.means.vocs[,-1], y = new.fem.means.vocs[,-1])
corrplot(ssex.corr.vocs, method = "circle", tl.col = "black", tl.cex = 0.8, tl.srt = 45)
```



