---
title: "Volatiles - EDA"
author: "Javiera Rudolph"
date: "5/17/2019"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
```

The original raw data excel file (LK_CompositeData_w_Partial_Flux_Final.xlsx) is fairly complicated. It has multiple sheets, references and calculations. The first sheet is the one that contains all of the data. So, I tried importing that one using the `readxl::read_xlsx` function but it failed to import. There must have been some formatting issue. I opted to use `read_csv` and transform the first sheet of the excel data into a csv file. I couldn't delete any of the morphological traits since it change the references for the volatile data. I just left it as is, removed all formating with colors, highlights, merged cells and removed the first explanatory cell. All of this was done by hand and not scripted. 


```{r}
rawdata <- read_csv("rawdata/20190517_JR_fromLKcompData.csv")
names(rawdata)
# select only the family, sample name and the volatiles
voldata <- rawdata[,c(1,2,54:234)]
names(voldata)
```

Some of the questions that arise from this dataset (which is only including volatiles and identifier, no sex is included):
 - What does the *Sample Name* stand for? And how many of each should we have. Some seem to be repeated numbers, but not all. 
 - What does the name of each volatile correspond to? What is the difference between **FLUX** and **(Conc)2**? Are they in any specific units?
