---
title: "EDA and creating a clean master data file"
output: 
  github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.width = 14, fig.height = 4)

```

```{r libraries}
library(tidyverse)
library(vegan)
library(plotly)
library(corrplot)
library(ggpubr)
library(Hmisc)

theme_set(theme_bw())
```


# Introduction
This document will contain pretty much everything we have done so far. This works as a draft reference of where we are and what we have done. 

# Data cleaning
Data cleaning is a really important section since lots of our functions will throw errors with NAs, zeroes or any typos. 
```{r}
# Read in the raw data
rawdata <- read.csv("rawdata/LK_master.csv", stringsAsFactors = FALSE)
```


## Trait data
To make things a little easier we are going to tackle the trait data first and then we will go on to clean VOCs. Some changes that we wish to accomplish here include changing the variable names to simpler/shorter ones without special characters. We would also like to remove empty rows at the end, manage the zeroes and NAs. There are some errors in measurements and typos that we need to address. 

```{r}

trait_raw <- rawdata[,1:19] %>% 
  rename(famid = `Fam..`,
         sampid = Sample.Name,
         ssex = Sample_Sex) %>% 
  drop_na(famid) %>% 
  mutate(sampid = str_replace_all(sampid, "\\(.*\\)", ""),
         sampid = str_trim(sampid, side = "both"),
         ssex = str_to_lower(as.character(ssex))) %>% 
  mutate_at(vars(starts_with("Leaf")), as.numeric) %>% 
  group_by(famid) %>% 
  arrange(sampid, .by_group = TRUE) %>% 
  ungroup

# Fix a data entry error:
# Check Github issue #5
trait_raw$Avg_21_days[which(trait_raw$Avg_21_days == 33)] <- 0.33

# Fix sex data entry error
# Check Github issue #5
trait_raw[which(trait_raw$sampid == "P_18_1_6_B"),]$ssex <- "m"
trait_raw <- trait_raw[-which(trait_raw$sampid == "P_6_6_20"),]
```

#### Leaf traits
First, we visually check for outliers or values that may not make sense.
```{r}
# Leaf data averaging -----------------------------------------------------

trait_raw[,17:19] %>% 
  gather(key = "trait", value = "value") %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```

There are no clear outliers or errors, so these look ok to average and we create a new data frame with the leaf data. In cases where data for only one of the clones was available, we kept that value. For cases where data was available for both clones we averaged it.  
```{r}
trait_raw[,c(2, 17:19),] %>% 
  group_by(sampid) %>% 
  summarise_at(vars(starts_with("Leaf")), mean, na.rm = TRUE) %>% 
  mutate_at(vars(starts_with("Leaf")), list(~ as.numeric(ifelse(. == "NaN", "NA", .)))) -> leaf_data

```

#### Growth and development traits
We do the same thing with these other traits and it is easy to identify three values that seem incorrect. These are in the area, circularity and perimeter rate.

```{r}
# Growth and Development traits -------------------------------------------

trait_raw[,10:16] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```

```{r}
outliers <- which(trait_raw$Avg_Area._Week_3>50)
trait_raw[outliers,]
```

```{r}
trait_raw$Avg_Area._Week_3[outliers] <- trait_raw$Avg_Area._Week_3[outliers] / 1000
trait_raw$Avg_Circularity_Week.3[outliers] <- trait_raw$Avg_Circularity_Week.3[outliers] / 1000

# check again 
trait_raw[,10:16] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```

These values had mistakes in the number of decimal places and included a typo making a value be negative. After adjusting these values, we check again for errors. The data looks ok now, so we can save the dataframe.
```{r}
negs <- which(trait_raw$Avg_Perimeter_Rate < 0)
trait_raw[negs,]

trait_raw$Avg_Perimeter_Rate[negs]
trait_raw$Avg_Perimeter_Rate[negs] <- trait_raw$Avg_Perimeter_Rate[negs] /-10

# Check visually
trait_raw[,10:16] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```


```{r}
trait_raw[,c(2, 10:16)] %>% 
  drop_na() %>% 
  distinct()-> gro_dev_data
```


#### Reproduction variable
We have different reproduction variables for males and females, so we can scale them and that way keep them under one variable only. This could be wrong, but I honestly don't know what else we could do.
```{r}
# Reproduction variable ---------------------------------------------------

trait_raw %>% 
  filter(ssex == "m") %>% 
  select(sampid, ssex, Avg_Male_Buds.Stem)%>% 
  drop_na() %>% 
  group_by(sampid) %>% 
  summarise(raw_av = mean(Avg_Male_Buds.Stem)) %>% 
  mutate(reprovar = scale(raw_av),
         ssex = "m")  -> male_reprovar

trait_raw %>% 
  select(sampid, ssex, Avg_Arch) %>% 
  filter(ssex == "f") %>% 
  drop_na() %>% 
  group_by(sampid) %>% 
  summarise(raw_av = mean(Avg_Arch)) %>% 
  mutate(reprovar = scale(raw_av),
         ssex = "f") -> fem_reprovar


reprovar <- bind_rows(male_reprovar, fem_reprovar)
reprovar %>% 
  ggplot(aes(x = reprovar, fill = ssex)) +
  geom_histogram()
```

### Clean trait data

We can now join all the smaller datasets and have a master traits data frame. These are not scaled so far and we can evaluate the changes of scaling them vs keeping them as they are.
```{r}
# Join trait data ---------------------------------------------------------

trait_raw %>% 
  select(famid, sampid, ssex) %>% 
  distinct() %>% 
  full_join(., reprovar) %>% 
  full_join(., gro_dev_data) %>%
  full_join(., leaf_data) %>% 
  set_names(c("famid", "sampid", "ssex", "raw_repro", "repro", "area_wk3",
              "perim_wk3", "circ_wk3", "perim_rate", "area_rate",
              "days21", "days_gam", "leaf_length", "leaf_area",
              "leaf_perim")) -> traits
```

#### Histograms raw trait data

```{r}
traits %>% 
  select(-c(famid, sampid, ssex, raw_repro)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")
```

#### Histograms transformed trait data
If we log transform the data and then scale it, the following figure shows the histograms for all the variables. Since we have variables in different units, we will use the scaled traits for future analyses and models.

```{r}
names(traits)
traits %>% 
  mutate_at(c(5:15), list(~ log10(. + 1))) %>% 
  mutate_at(c(5:15), scale) -> scaled_traits

scaled_traits %>% 
  select(-c(famid, sampid, ssex, raw_repro)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#7e9bcc") +
  facet_wrap( ~ trait, scales = "free")
```


## Volatile Organic Compounds

Using the same raw data csv file, we are going to select only the identifiers and the VOC data. 

```{r}
# Just keep volatile data and identifiers
# Also, remove the NAs, there's a lot at the end
# Remove '(2)' notation from sample ids

#names(rawdata)
voc_raw <- rawdata[,c(1,2,4,5, 20:109)] %>% 
  rename(famid = `Fam..`,
         sampid = Sample.Name,
         ssex = Sample_Sex) %>% 
  drop_na() %>% 
  mutate(sampid = str_replace_all(sampid, "\\(.*\\)", ""),
         sampid = str_trim(sampid, side = "both"))

#' Shorten voc names
oldnames <- names(voc_raw)
names(voc_raw)[5:94] <- stringr::str_trunc(oldnames[5:94], width = 6, side = "right", ellipsis = "")

```

There was an issue with the notation in one of the compounds, so we check that and fix the notation so that R can recognize the values as numbers.
```{r}
#str(voc_raw)
#' m111.0 is a character
#voc_raw$m111.0
#' The notation is different, using a capital E
voc_raw$m111.0 <- as.numeric(str_replace(voc_raw$m111.0, "E", "e"))
```

### May 30 samples
Leslie mentioned that data associated to May 30 was really off, so we are going to check that and probably remove it if the values are not similar to the rest of the dataset. With the figure it is obvious that we need to remove the May 30 data. With these boxplots we see that there are some negative values. These come from subtracting the noise, so any negative values should be zero.

```{r message=FALSE}
voc_raw %>% 
  dplyr::select(-c(famid, sampid, ssex)) %>% 
  gather(., key = "compound", value = "value", -Date.PTR) %>% 
  ggplot(aes(x = Date.PTR, y = value)) +
  geom_boxplot() -> withMay30

voc_raw %>% 
  filter(Date.PTR != "30-May") %>%
  mutate_if(is.numeric, list(~ ifelse(. < 0, 0, .))) %>% 
  arrange(sampid, .by_group = TRUE) %>% 
  dplyr::select(-c(famid, sampid, ssex)) %>% 
  gather(., key = "compound", value = "value", -Date.PTR) %>% 
  ggplot(aes(x = Date.PTR, y = value)) +
  geom_boxplot() ->woutMay30

ggarrange(withMay30, woutMay30, nrow = 2)

```

### Using a 10% threshold
We are only going to keep the compounds for which at least 10% of the samples have a value greater than zero. According to Leslie, other sources use up to 40% of a threshold, however, we are interested in sexual dimorphism. That means that if there is a difference in compound production associated to the sex, then you automatically have only a 50% chance of producing that compound. Therefore, a 10 or 20% threshold is what would work for this project. 

```{r}
voc_raw %>% 
  filter(Date.PTR != "30-May") %>%
  mutate_if(is.numeric, list(~ ifelse(. < 0, 0, .))) %>% 
  arrange(sampid, .by_group = TRUE) %>% 
  dplyr::select(-Date.PTR) -> vocs

# Create a dataset with only the compounds for which at least 10% of the observations show it.
vocs %>% 
  select(-famid, -sampid, -ssex) %>% 
  mutate_if(is.numeric, function(x) ifelse(x > 0, 1, 0)) %>% 
  colSums() %>% 
  data.frame() %>% 
  rownames_to_column() %>%
  setNames(., c("voc", "count")) %>% 
  mutate(prcnt = count/nrow(vocs)) %>% 
  filter(prcnt >= 0.1) -> voc_filter

vocs %>% 
  select(famid, sampid, ssex, c(voc_filter$voc)) %>% 
  group_by(sampid) %>%
  mutate_at(vars(starts_with("m")), max) %>% 
  ungroup () %>% 
  distinct() %>% 
  mutate(ssex = str_to_lower(as.character(ssex))) -> voc_data

```

### VOC Clustering

For us to cluster the VOC data, we need to look at it first. 

```{r}
long_voc_data <- voc_data %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))

```

#### VOC data exploration
We know we are working with very small numbers and this influences how we will manage the data and what sort of clustering we can do for it. With a quick check, we see that the numbers are very small and that different compounds have very different ranges as well.
```{r}
long_voc_data %>% 
  ggplot(aes(x = voc, y = conc)) + 
  geom_point() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90))
```

After looking at the data, the first thought is to log transform it. To check for any measurement outliers we might want to get rid of the zeroes and transform the data to identify potential outliers.

```{r}
long_voc_data %>% 
  filter(conc > 0) %>%
  ggplot(aes(x = voc, y = conc)) + 
  geom_boxplot() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10()
```

With the information, we can clean the data again and remove these outliers. Since the values are so small and below the machine's detection threshold, they are classified as zero.
```{r}
# Making these outliers be zero
voc_data[voc_data < 10e-15] <- 0

long_voc_data <- voc_data %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))


```

```{r eval = FALSE}
long_voc_data %>% 
  ggplot(aes(x = voc, y = conc)) + 
  geom_boxplot() +
  theme_bw() + 
  lims(y = c(1e-13, 1e-05)) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10()
```
#### Transformation and scaling

Based on PTR machines detection threshold, any values smaller than `1e-12` are equivalent to zero. Then we will standardize the data.
```{r}
log_vocs <- voc_data %>% 
  mutate_at(vars(starts_with("m")), list(~ log10(. + 1e-12)))

st_log_vocs <- log_vocs %>% 
  mutate_at(vars(starts_with("m")), scale)

long_vocs <- st_log_vocs %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))

```

This is how the transformed data looks like now, with a visual differentiation between males and females. Recall that the data for VOCs here has been log transfomed and scaled. 
```{r fig.width=24, fig.height=8}
ggarrange(long_vocs %>% 
  ggplot(aes(x = voc, y = conc)) +
  geom_boxplot(notch = TRUE, alpha = 0.2) +
  labs(x = "VOC ID",
       y = "Log transformed and scaled concentration") +
  theme(axis.text.x = element_text(angle = 90)),
  long_vocs %>% 
  ggplot(aes(x = voc, y = conc, fill = ssex)) +
  geom_boxplot(notch = TRUE, alpha = 0.2) +
  labs(x = "VOC ID",
       y = "Log transformed and scaled concentration") +
  theme(axis.text.x = element_text(angle = 90)), nrow = 2)
  
```

#### Correlations
The point of running these correlations is to get an idea of which compounds might show up together or the opposite. We are trying to reduce our dataset and avoid inputing 75 dependent variables into a model with two explanatory variables. From the correlation matrix we can see that some compounds tend to be found in clusters. We can use this information to group the compounds and get a reduced dataset.

```{r fig.width=10, fig.height=10}

cor_data <- as.matrix(st_log_vocs[,4:78])

cor_res <- rcorr(x = cor_data, type = "pearson")

saveRDS(cor_res, "cleandata/data_for_figs/voc_correlation_data.RDS")
corrplot(cor_res$r, p.mat = cor_res$P, sig.level = 0.001, insig = "pch", tl.col = "black", tl.srt = 60, tl.cex = .4, order = "hclust", pch.cex = 0.8, pch.col = "#43484f", addrect = 17)
```


#### Clustering

Haven't found a good reasoning to come up with a specific number of clusters. However, by making these clusters using the correlation matrix, we can see how variables (in this case compounds) are grouped together.

```{r}
d <- as.dist(1-cor_res$r)
hr <- hclust(d)
#names(hr)
plot(hr, hang = -1)
rect.hclust(hr, h = max(hr$height)/2)
```

```{r eval = FALSE, echo = FALSE}
# This wa swith arbitrary number of clusters and also not sure about using dist vs as.dist(1-data). Since we are using correlation, I'm incluined to the second option.
voc_dist <- hclust(dist(cor_res$r))
plot(voc_dist, hang = -1)
rect.hclust(voc_dist, k =15)

voc_clusters <- data.frame(clust = cutree(voc_dist, k = 15)) %>% 
  rownames_to_column(var = "voc")
```

Using the clustering to create a new data frame
**Making clusters using the original raw data and then transforming and scaling, after the clusters are formed**  

```{r}
# This uses the original, raw untransformed data to get the means by cluster
voc_clusters <- data.frame(clust = cutree(hr, h = max(hr$height)/2)) %>% 
  rownames_to_column(var = "voc") 


long_voc_data %>% 
  left_join(voc_clusters) %>% 
  group_by(sampid, clust, famid, ssex) %>% 
  summarise(voc_value = mean(conc)) %>% 
  mutate(voc_clust = str_pad(clust, width = 2, side = "left", pad = 0),
         voc_clust = paste0("clust_", voc_clust)) %>% 
  ungroup() %>% 
  select(-clust) -> clustered_long_vocs

clustered_long_vocs %>% 
  spread(key = voc_clust, value = voc_value) -> clustered_voc_data


clustered_log_vocs <- clustered_voc_data %>% 
  mutate_at(vars(starts_with("c")), list(~ log10(. + 1e-12)))

clustered_st_log_vocs <- clustered_log_vocs %>% 
  mutate_at(vars(starts_with("c")), scale)

clustered_long_vocs <- clustered_st_log_vocs %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))

clustered_long_vocs %>% 
  ggplot(aes(x = voc, y = conc, fill = ssex)) +
  geom_boxplot(notch = TRUE, alpha = 0.2) +
  labs(x = "VOC ID",
       y = "Log transformed and scaled concentration") +
  theme(axis.text.x = element_text(angle = 90))

```
**Making the clusters with the data that was already transformed and scaled**
I think it makes more sense to use this one, since this is the data we actually used to create the clusters. Our correlations and everything else is based on this transformed data, so this is probably what we should use to create the new dataset of clustered VOCs.

```{r}

# This one gets the clustered data by getting the mean using the transformed and scaled data.
voc_clusters <- data.frame(clust = cutree(hr, h = max(hr$height)/2)) %>% 
  rownames_to_column(var = "voc") 


long_vocs %>% 
  left_join(voc_clusters) %>% 
  group_by(sampid, clust, famid, ssex) %>% 
  summarise(voc_value = mean(conc)) %>% 
  mutate(voc_clust = str_pad(clust, width = 2, side = "left", pad = 0),
         voc_clust = paste0("clust_", voc_clust)) %>% 
  ungroup() %>% 
  select(-clust) -> clustered_long_vocs2

clustered_long_vocs2 %>% 
  spread(key = voc_clust, value = voc_value) -> clustered_voc_data2

clustered_long_vocs2 %>% 
  ggplot(aes(x = voc_clust, y = voc_value, fill = ssex)) +
  geom_boxplot(notch = TRUE, alpha = 0.2) +
  labs(x = "VOC ID",
       y = "Log transformed and scaled concentration") +
  theme(axis.text.x = element_text(angle = 90))
```

These are the clusters and the compounds in each cluster
```{r echo = FALSE}
voc_clusters %>% 
  dplyr::select(clust, voc) %>% 
  rename(Cluster = clust, Compound = voc) %>% 
  arrange(Cluster) %>% 
  knitr::kable()
```



### Master dataframe
This will create the master data frame that includes the lifehistory traits and the VOC clustered data. All of these are already log transformed and scaled.

```{r}
clean_master <- full_join(scaled_traits, clustered_voc_data2) %>% 
  mutate_at(c(5:32), as.numeric)

saveRDS(clean_master, "cleandata/clean_master.RDS")
```

