---
title: "Evolution Talk: figures"
output: 
  github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.width = 12, fig.height = 4)

```

```{r libraries}
library(tidyverse)
library(vegan)
library(corrplot)
library(ggpubr)
library(Hmisc)

theme_set(theme_bw())
```

```{r}
master <- readRDS("cleandata/clean_master.RDS")
```

# VOCs data
## Correlation matrix of all the vocs with the clusters

We wanted to see if some compounds could be potentially related. We ran a correlation. The grey 'x' in the figure shows insignificant correlations. We used a hierarchical clustering algorithm to group the VOCs in clusters.

```{r fig.width=10, fig.height=10}
cor_res <- readRDS("cleandata/data_for_figs/voc_correlation_data.RDS")
corrplot(cor_res$r, p.mat = cor_res$P, sig.level = 0.001, insig = "pch", tl.col = "black", tl.srt = 60, tl.cex = .4, order = "hclust", pch.cex = 0.8, pch.col = "#43484f", addrect = 17)
```

### MANOVA for vocs
```{r}
vocs <- master[,c(1, 3, 16:32)] %>% 
  drop_na()

vocs.manova <- manova(cbind(clust_01, clust_02, clust_03, clust_04,
                            clust_05, clust_06, clust_07, clust_08,
                            clust_09, clust_10, clust_11, clust_12,
                            clust_13, clust_14, clust_15, clust_16,
                            clust_17) ~ famid + ssex + famid:ssex, data = vocs)

sum.vocs.manova <- summary(vocs.manova)
sum.vocs.manova
summary.aov(vocs.manova)
```


## Cluster boxplots for males and females
From the MANOVA results, all of these clusters show sexual dimorphism.

```{r}
vocs %>% 
  gather(key = "voc_clust", value = "voc_value", -c(famid, ssex)) %>% 
  ggplot(aes(x = voc_clust, y = voc_value, fill = ssex)) +
  geom_boxplot(notch = TRUE, alpha = 0.2) +
  labs(x = "VOC ID",
       y = "Log transformed and scaled concentration") +
  theme(axis.text.x = element_text(angle = 90))
  
```



# Correlation matrices for all traits
## Side by side males and females
## males on one axis, females on the other
### Highlight the diagonal with transparencies
## Subtracted matrix

# Relationship between VOC clusters and the life history traits such as leafs






















