---
title: "Evolution Talk: Figures"
author: "Javiera Rudolph"
date: "6/13/2019"
output:
  github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

```

```{r libraries}
library(tidyverse)
library(vegan)
library(corrplot)
library(ggpubr)
library(Hmisc)

theme_set(theme_bw())
```

```{r}
master <- readRDS("cleandata/clean_master.RDS") %>% 
  dplyr::select(-raw_repro)
```

# VOCs data
## Correlation matrix of all the vocs with the clusters

We wanted to see if some compounds could be potentially related. We ran a correlation. The grey 'x' in the figure shows insignificant correlations. We used a hierarchical clustering algorithm to group the VOCs in clusters.

```{r fig.width=10, fig.height=10}
cor_res <- readRDS("cleandata/data_for_figs/voc_correlation_data.RDS")
corrplot(cor_res$r, p.mat = cor_res$P, sig.level = 0.001, insig = "pch", tl.col = "black", tl.srt = 60, tl.cex = .4, order = "hclust", pch.cex = 0.8, pch.col = "#43484f", addrect = 17)
```

### MANOVA for vocs
```{r}
vocs <- master[,c(1, 3, 15:31)] %>% 
  drop_na()

vocs.manova <- manova(cbind(clust_01, clust_02, clust_03, clust_04,
                            clust_05, clust_06, clust_07, clust_08,
                            clust_09, clust_10, clust_11, clust_12,
                            clust_13, clust_14, clust_15, clust_16,
                            clust_17) ~ famid + ssex + famid:ssex, data = vocs)

sum.vocs.manova <- summary(vocs.manova)
sum.vocs.manova
summary.aov(vocs.manova)
```


## Cluster boxplots for males and females
From the MANOVA results, all of these clusters show sexual dimorphism.

```{r fig.width=12, fig.height=5}
vocs %>% 
  gather(key = "voc_clust", value = "voc_value", -c(famid, ssex)) %>% 
  ggplot(aes(x = voc_clust, y = voc_value, fill = ssex)) +
  geom_boxplot(notch = TRUE, alpha = 0.2) +
  labs(x = "VOC ID",
       y = "Log transformed and scaled concentration") +
  theme(axis.text.x = element_text(angle = 90))

```



# Correlation matrices for all traits
## Side by side males and females

```{r}
# Get the data frames for both males and females
master %>% 
  filter(ssex == "m") %>% 
  group_by(famid) %>% 
  mutate_at(c(4:31), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex)) %>% 
  distinct() -> male_data

master %>% 
  filter(ssex == "f") %>% 
  group_by(famid) %>% 
  mutate_at(c(4:31), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex)) %>% 
  distinct() %>% 
  filter(famid %in% male_data$famid) -> fem_data

```


```{r}
# Males - correlation
cor_males <- cor(male_data[,2:29], use = "na.or.complete")
# Females - correlation
cor_fem <- cor(fem_data[,2:29], use = "na.or.complete")
```


```{r fig.width=9, fig.height=9}
corrplot(cor_males, title = "Males", type = "upper", tl.col = "black", tl.srt = 60, tl.cex = .8, mar = c(0,0,2,0))

corrplot(cor_fem, title = "Females", type = "upper", tl.col = "black", tl.srt = 60, tl.cex = .8, mar = c(0,0,2,0))

```

## males on one axis, females on the other
```{r fig.width=9, fig.height=9}
# Make sure you have the same number of families to get this data
# Seems like they both have 44

sex_cor <- cor(male_data[,2:29], fem_data[,2:29], use = "na.or.complete")
corrplot(sex_cor, title = "Males on X, Females on Y", tl.col = "black", tl.srt = 60, tl.cex = .8, mar = c(0,0,2,0))


```



### Highlight the diagonal with transparencies
To get fancy with these plots we need to switch over to ggplot.

#### Option 1
 
```{r}

order_variables <- c(sex_cor %>% 
  as.data.frame() %>% 
  names())


sex_cor %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "males") %>% 
  mutate(males = factor(males, levels = order_variables)) %>% 
  gather(key = "females", value = "trait", -males) %>% 
  mutate(females = factor(females, levels = order_variables)) %>% 
  mutate(diag = males == females) %>%
  ggplot(aes(x = males, y = females, fill = trait, color = diag)) +
  geom_tile() +
  scale_fill_gradient2(low = "#0000FF", mid = "#FFFFFF", high ="#FF0000") +
  scale_color_manual(values = c("white", "black")) +
  guides(color = 'none') +
  theme(axis.text.x = element_text(angle = 90))

```

#### Option 2
I think the transparencies are misleading because it makes you think that there is no correlation... 

```{r}

order_variables <- c(sex_cor %>% 
  as.data.frame() %>% 
  names())


sex_cor %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "males") %>% 
  mutate(males = factor(males, levels = order_variables)) %>% 
  gather(key = "females", value = "trait", -males) %>% 
  mutate(females = factor(females, levels = order_variables)) %>% 
  mutate(diag = as.numeric(males == females)) %>% 
  ggplot(aes(x = males, y = females, fill = trait, alpha = diag)) +
  geom_tile() +
  scale_fill_gradient2(low = "#0000FF", mid = "#FFFFFF", high ="#FF0000") +
  guides(alpha = 'none') +
  theme(axis.text.x = element_text(angle = 90))

```

## Subtracted matrix

```{r}
# If you want to consider the males subtracted from female matrices
# From the correlation matrices

minus_mat <- cor_males - cor_fem
corrplot(minus_mat, type = "upper") # you can do this with corrplot

```

Also can do it with ggplot
```{r}
order_variables <- c(minus_mat %>% 
  as.data.frame() %>% 
  names())


minus_mat %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "males") %>% 
  mutate(males = factor(males, levels = order_variables)) %>% 
  gather(key = "females", value = "trait", -males) %>% 
  mutate(females = factor(females, levels = order_variables)) %>% 
  ggplot(aes(x = males, y = females, fill = trait)) +
  geom_tile() +
  scale_fill_viridis_c() +
  #scale_fill_gradient2(low = "#0000FF", mid = "#FFFFFF", high ="#FF0000") +
  theme(axis.text.x = element_text(angle = 90))
```


# Relationship between VOC clusters and the life history traits such as leafs


