---
title: "G matrix - brute force"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(corrplot)
```


```{r}
traits <- readRDS("cleandata/traits_clean.RDS")
vocs <- readRDS("cleandata/clustered_voc_data.RDS")
str(vocs)


vocs %>% 
  mutate_at(vars(starts_with("c")), list(~ log10(. + 1))) %>% 
  mutate_at(vars(starts_with("c")), scale) -> vocs


theme_set(theme_bw())
```


# Data transformation
There are NAs in the data since some traits weren't measured. These are a problem when we do a log transformation. In my opinion, I wouldn't do a log transformation, just scale the variables. Specially for the days21 variable, we have lots of zeroes and will probably run into trouble there.

```{r}

traits %>% 
  select(-c(famid, sampid, ssex, raw_repro)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")

```

If we scale the variables, they all look normal, except for days 21 and days gam. Which makes sense because these are different variables, they are not measurements, they are time to an event.

```{r}
traits %>% 
  select(-c(famid, sampid, ssex, raw_repro)) %>%
  #mutate_all(log10) %>% 
  mutate_all(scale) %>% 
  gather(key = "trait", value = "standardized_values") %>% 
  ggplot(aes(x = standardized_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#87afac") +
  facet_wrap( ~ trait, scales = "free")
```

#Scaling volatile data
```{r}
vocs %>% 
  select(-c(famid, sampid, ssex)) %>%
  mutate_all(scale) %>% 
  gather(key = "trait", value = "standardized_values") %>% 
  ggplot(aes(x = standardized_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8766a8") +
  facet_wrap( ~ trait, scales = "free")

```

#Scaling traits 
```{r}
traits %>% 
  mutate_at(c(5:15), scale) -> scaled_traits

#Combining volatile and life history data into one master file. Scaling can occur prior to combining because the function works row by row
master <- full_join(x = scaled_traits, y = vocs)
```

# JUST TRAITS DATA
```{r}
mynamestheme <- theme(plot.title = element_text(family = "Arial", face = "bold", size = (20)), 
                 legend.title = element_text(colour = "black",  face = "bold.italic", family = "Arial"), 
                 legend.text = element_text(face = "italic", colour="black",family = "Arial"), 
                  axis.title = element_text(family = "Arial", size = (13), colour = "Black"),
                  axis.text = element_text(family = "Arial", colour = "black", size = (10)))
  

scaled_traits %>% 
  filter(ssex == "f") %>% 
  group_by(famid) %>% 
  mutate_at(c(5:15), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex, raw_repro)) %>% 
  distinct() %>% 
  select(-famid) %>% 
  cov(use = "na.or.complete") %>% 
  as.data.frame() -> cov_mat_fems

cov_mat_fems %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  mynamestheme +
  ggtitle("Female life history trait covariance matrix")+
  scale_fill_viridis_c(option = "magma") +
  theme(axis.text.x = element_text(angle = 90))
```

Correlation matrix females life history traits
```{r}
cov_mat_fems <- as.matrix(cov_mat_fems)
cor_mat_fems <- (cov2cor(cov_mat_fems))
females_life <- corrplot(cor_mat_fems, title= "Correlation matrix females life history traits", tl.col = "black", mar=c(0,2,1,1))


```

MALE TRAIT COVARIANCE MATRIX
```{r}
scaled_traits %>% 
  filter(ssex == "m") %>% 
  group_by(famid) %>% 
  mutate_at(c(5:15), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex, raw_repro)) %>% 
  distinct() %>% 
  select(-famid) %>% 
  cov(use = "na.or.complete") %>% 
  as.data.frame() -> cov_mat_male

cov_mat_male %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c() +
  mynamestheme+
  ggtitle("Male trait data covariance matrix")+
  theme(axis.text.x = element_text(angle = 90))
```

Correlation matrix males life history traits
```{r}
cov_mat_male <- as.matrix(cov_mat_male)
cor_mat_male <- (cov2cor(cov_mat_male))
Males_life <- corrplot(cor_mat_male, title= "Correlation matrix males life history traits", tl.col = "black", mar=c(0,2,1,1))
```

#ALL TRAIT DATA COVARIANCE MATRIX but not seperated by sex.
```{r}
scaled_traits %>% 
  group_by(famid) %>% 
  mutate_at(c(5:15), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex, raw_repro)) %>% 
  distinct() %>% 
  select(-famid) %>% 
  cov(use = "na.or.complete") %>% 
  as.data.frame() -> cov_mat_all

cov_mat_all %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma") +
  mynamestheme+
  ggtitle("All trait data covariance. Sexes not seperated out")+
  labs(x= "Family means all sexes", y="Family means all sexes")+
  theme(axis.text.x = element_text(angle = 90))

```

Correlation matrix for all life history traits
```{r}

cov_mat_all <- as.matrix(cov_mat_all)
cor_mat_all <- (cov2cor(cov_mat_all))
Males_life <- corrplot(cor_mat_all ,title= "Correlation matrix all life history traits. Not seperate by sex", tl.col = "black", mar=c(0,2,1,1))
```

#All traits by filtering with only families shared between males and females. Plotted males vs females
All traits filtered males v females covariance
```{r}


scaled_traits %>% 
  filter(ssex == "m") %>% 
  group_by(famid) %>% 
  mutate_at(c(5:15), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex, raw_repro)) %>% 
  distinct() -> male_data

scaled_traits %>% 
  filter(ssex == "f") %>% 
  group_by(famid) %>% 
  mutate_at(c(5:15), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex, raw_repro)) %>% 
  distinct() %>% 
  filter(famid %in% male_data$famid) -> fem_data

# Use only the families that have both males and females
test_cov <- cov(male_data[,2:12], fem_data[,2:12], use = "pairwise.complete.obs")
test_cor <- cor(male_data[,2:12], fem_data[,2:12])


test_cov %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  labs(x="Male", y="Female")+
  ggtitle("Covariance matrix of life histroy data for males vs females")+
  mynamestheme+
  theme(axis.text.x = element_text(angle = 90))

```

All traits filtered males v females correlation
```{r}

test_cor %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
labs(x="Male", y="Female")+
  ggtitle("Correlation matrix of life histroy data for males vs females")+
  mynamestheme+
  theme(axis.text.x = element_text(angle = 90))
```


# Just VOCs
Female volatiles covariance plot

```{r}
vocs %>% 
  filter(ssex == "f") %>% 
  group_by(famid) %>% 
  mutate_at(c(5:15), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex)) %>% 
  distinct() %>% 
  select(-famid) %>% 
  cov(use = "na.or.complete") %>% 
  as.data.frame() -> cov_mat_fems

cov_mat_fems %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c(option = "magma") +
  mynamestheme+
  ggtitle("Female volatiles covariance plot")+
  theme(axis.text.x = element_text(angle = 90))

```

Female volatiles correlation plot
```{r}
cov_mat_fems <- as.matrix(cov_mat_fems)
cor_mat_fems <- (cov2cor(cov_mat_fems))
females_voc <- corrplot(cor_mat_fems,title= "Correlation matrix female volatiles", tl.col = "black", mar=c(0,2,1,1))
```

Male volatile covariance matrix
```{r}
vocs %>% 
  filter(ssex == "m") %>% 
  group_by(famid) %>% 
  mutate_at(c(5:15), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex)) %>% 
  distinct() %>% 
  select(-famid) %>% 
  cov(use = "na.or.complete") %>% 
  as.data.frame() -> cov_mat_male

cov_mat_male %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c() +
  ggtitle("Male volatile covariance matrix")+
  mynamestheme+
  theme(axis.text.x = element_text(angle = 90))

```

Male volatile correlation matrix
```{r}

cov_mat_male <- as.matrix(cov_mat_male)
cor_mat_male <- (cov2cor(cov_mat_male))
males_voc <- corrplot(cor_mat_male, title= "Correlation matrix male volatiles", tl.col = "black", mar=c(0,2,1,1))
```

#All data not female vs male
```{r}
vocs %>% 
  group_by(famid) %>% 
  mutate_at(c(5:15), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex)) %>% 
  distinct() %>% 
  select(-famid) %>% 
  cov(use = "na.or.complete") %>% 
  as.data.frame() -> cov_mat_all

cov_mat_all %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma") +
  mynamestheme+
  ggtitle("Covariance matrix with all volatile data but not female vs male")+
  theme(axis.text.x = element_text(angle = 90))


```

Male volatile correlation matrix
```{r}
cov_mat_all <- as.matrix(cov_mat_all)
cor_mat_all <- (cov2cor(cov_mat_all))
all_VOC <- corrplot(cor_mat_all,title= "Correlation matrix all volatile data but not male vs female", tl.col = "black", mar=c(0,2,1,1))
```

#All vocs filtered by those with Male and female families. Plot M v F
Male vs female covariance
```{r, eval = FALSE}
vocs %>% 
  filter(ssex == "m") %>% 
  group_by(famid) %>% 
  mutate_at(c(4:18), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex)) %>% 
  distinct() -> male_data

vocs %>% 
  filter(ssex == "f") %>% 
  group_by(famid) %>% 
  mutate_at(c(4:18), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex)) %>% 
  distinct() %>% 
  filter(famid %in% male_data$famid) -> fem_data

# Use only the families that have both males and females

fem_data %>%
  filter(famid %in% male_data$famid) -> fem_data

test_cov <- cov(male_data[,2:16], fem_data[,2:16], use = "na.or.complete")
test_cor <- cor(male_data[,2:16], fem_data[,2:16], use = "na.or.complete")

#Male vs female covariance
test_cov %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  ggtitle("Covariance matrix for volatiles data with males vs females")+
  mynamestheme+
  labs(x="Males", y="Females")+
  theme(axis.text.x = element_text(angle = 90))


```

Male vs female correlation
```{r}
test_cor %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  ggtitle("Correlation matrix of volatile data males vs females")+
  mynamestheme+
  labs(x="Males", y="Females")+
  theme(axis.text.x = element_text(angle = 90))
```

# Combined life history and volatile data
Female all covariance
```{r}
#Female data combined
master %>% 
  filter(ssex == "f") %>% 
  group_by(famid) %>% 
  mutate_at(c(5:30), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex, raw_repro)) %>% 
  distinct() %>% 
  select(-famid) %>% 
  cov(use = "na.or.complete") %>% 
  as.data.frame() -> cov_mat_fems

#Female all covariance
cov_mat_fems %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  ggtitle("Covariance matrix of all traits in females")+
  mynamestheme+
  scale_fill_viridis_c(option = "magma") +
  theme(axis.text.x = element_text(angle = 90))


```

Female all correlation
```{r}

cov_mat_all <- as.matrix(cov_mat_all)
cor_mat_all <- (cov2cor(cov_mat_all))
all_VOC <- corrplot(cor_mat_all)

```

#All data for males
```{r}
master %>% 
  filter(ssex == "m") %>% 
  group_by(famid) %>% 
  mutate_at(c(5:30), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex, raw_repro)) %>% 
  distinct() %>% 
  select(-famid) %>% 
  cov(use = "na.or.complete") %>% 
  as.data.frame() -> cov_mat_male

#Covariance matrix for males for all data
cov_mat_male %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  ggtitle("Covariance matrix of all traits in males")+
  scale_fill_viridis_c() +
  theme(axis.text.x = element_text(angle = 90))

```

male all correlation
```{r}
#
cov_mat_male <- as.matrix(cov_mat_male)
cor_mat_male <- (cov2cor(cov_mat_male))
all_male <- corrplot(cor_mat_male, title= "Correlation matrix all male traits", tl.col = "black", mar=c(0,2,1,1))
```


#All data but not filtered by sex
All data covariance
```{r}
#All data but not filtered by sex
master %>% 
  group_by(famid) %>% 
  mutate_at(c(5:30), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex, raw_repro)) %>% 
  distinct() %>% 
  select(-famid) %>% 
  cov(use = "na.or.complete") %>% 
  as.data.frame() -> cov_mat_all

#All data covariance
cov_mat_all %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma") +
  theme(axis.text.x = element_text(angle = 90))


```

all data correlation
```{r}

cov_mat_all <- as.matrix(cov_mat_all)
cor_mat_all <- (cov2cor(cov_mat_all))
all_alle <- corrplot(cor_mat_all)
```

#ALl data filtered by males v females
Covariance for males v female for all data
```{r eval = FALSE}


master %>% 
  filter(ssex == "m") %>% 
  group_by(famid) %>% 
  mutate_at(c(5:30), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex, raw_repro)) %>% 
  distinct() -> male_data

master %>% 
  filter(ssex == "f") %>% 
  group_by(famid) %>% 
  mutate_at(c(5:30), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(-c(sampid, ssex, raw_repro)) %>% 
  distinct() %>% 
  filter(famid %in% male_data$famid) -> fem_data

# Use only the families that have both males and females
#Covariance for males v female for all data
test_cov <- cov(male_data[,2:12], fem_data[,2:12], use = "na.or.complete")
test_cor <- cor(male_data[,2:12], fem_data[,2:12], use = "na.or.complete")

test_cov %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  ggtitle("Covariance matrix of all data in males v females")+
  mynamestheme+
  labs(x="Males", y="Females")+
  theme(axis.text.x = element_text(angle = 90))


```

Correlation for all data for males vs females
```{r}
#Correlation for all data for males vs females
test_cor %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "x") %>% 
  gather(key = "y", value = "value", -x) %>% 
  ggplot(aes(x = x, y = y, fill = value)) +
  geom_tile() +
  ggtitle("Correlation matrix of all data in males v femles")+
  mynamestheme+
  labs(x="Males", y="Females")+
  theme(axis.text.x = element_text(angle = 90))
```




