---
title: "Growth_Development_Traits_cleaning_data"
author: "Leslie Kollar"
date: "11/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 14, fig.height = 6)
```

## Data cleaning for life history traits
```{r libraries}
library(tidyverse)
library(vegan)
library(plotly)
library(corrplot)
library(ggpubr)
library(ggmosaic)
library(lme4)
library(lsmeans)
library(insight)
```


```{r dataload}
setwd("/Users/lesliekollar/Desktop/Moss_matrix/")
data_traits <- read.csv("rawdata/Data_Growth_experiment_summer_2017_not_all.csv")
```

```{r}
#Lots of DIV/0
data_traits$Week_3_Circularity <- as.numeric(str_replace(data_traits$Week_3_Circularity, "#DIV/0!", "NA"))
data_traits$Days_until_gametophores <- as.numeric(str_replace(data_traits$Days_until_gametophores, "#DIV/0!", "NA"))
data_traits$Total_Gametophores <- as.numeric(str_replace(data_traits$Total_Gametophores,"#DIV/0!", "NA"))
data_traits$Week_3_area <- as.numeric(str_replace(data_traits$Week_3_area, "#DIV/0!", "NA"))
data_traits$Week_3_perimeter <- as.numeric(str_replace(data_traits$Week_3_perimete, "#DIV/0!", "NA"))

#Filtering out rows with NAs because scale doees not like NA
data_traits %>% 
   filter_all(any_vars(is.na(.)))

data_traits[is.na(data_traits)] <- 0
```

```{r}
# Change variable names and fix sample ID names with spaces, or parentheses

trait <- data_traits %>% 
  drop_na(Family) %>% 
  mutate(Sample = str_replace_all(Sample, "\\(.*\\)", ""),
         Sample = str_trim(Sample, side = "both"),
         Sex = str_to_lower(as.character(Sex))) %>% # Sex on lower case only
  mutate_at(vars(starts_with("Week_3_area")), as.numeric) %>% # Make them all numeric 
  group_by(Family) %>% 
  arrange(Sex, .by_group = TRUE) %>% # Order by family number
  ungroup

trait$Week_3_Circularity <- as.numeric(trait$Week_3_Circularity)



```

## Checking for outliers and removing them
```{r}
#Visually check for outliers or values that do not make sense for area
trait[,5] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()

#Remove area week 3 outlier
outliers <- which(trait$Week_3_area>50)
trait[outliers,]

trait$Week_3_area[outliers] <- (trait$Week_3_area[outliers] / 1000)
                                
#Visually check for outliers or values that do not make sense for perimeter
trait[,6] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()

#Remove perimeter week 3 outlier
outliers <- which(trait$Week_3_perimeter>30)
trait[outliers,]

trait$Week_3_perimeter[outliers] <- (trait$Week_3_perimeter[outliers] / 1000)      


#Remove circularity week 3 outlier
trait[,7] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()

outliers <- which(trait$Week_3_Circularity>10)
trait[outliers,]

trait$Week_3_Circularity[outliers] <- (trait$Week_3_Circularity[outliers] / 1000) 
                                     
#Remove Days Until Gam--- really are those outliers though?
trait[,8] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()

outliers_Gam <- which(trait$Days_until_gametophores>25)
trait[outliers_Gam,]

trait$Days_until_gametophores[outliers_Gam] <- (trait$Days_until_gametophores[outliers_Gam] == 0) 


#Remove Total--- I dont think these are outliers either...
trait[,9] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()

#Double Check
trait[,5:6] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()

trait[,7] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()

trait[,c(1, 5:9)] %>% 
  drop_na() %>% 
  distinct()-> gro_dev_data
                                                                     
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Histogram of raw data
```{r}

#I think we can use the raw data for these traits moving forward. 
trait %>% 
  select(-c(Family, Sample, Sex, Plate)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")
#Remove Sex=x
trait<-trait[!(trait$Sex=="x"),]

#saveRDS( , "cleandata/scaled_traits_Nov_26")
```

```{r}
#Scaling the traits and not log transforming them
names(trait)
trait %>% 
  #mutate_at(c(5:9), list(~ log10(. + 1))) %>% 
  mutate_at(c(5:9), scale) -> scaled_traits

scaled_traits %>% 
  select(-c(Family, Sample, Sex, Plate)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#7e9bcc") +
  facet_wrap( ~ trait, scales = "free")

scaled_traits <- scaled_traits[,1:9]

```




```{r}
saveRDS(scaled_traits, "cleandata/Final_growth_dev_traits.RDS")
```

##Running ordination on growth and development traits
```{r}
#Run PCA
scaled_pca <- rda(scaled_traits[,5:9])



# Figure for log transformed data
biplot(scaled_pca, display = c("sites", 
                   "species"),
       type = c("text",
                "points"))
ssex.colors <- levels(factor(trait$Sex))
ordihull(scaled_pca, group = factor(trait$Sex),
         col = c("green","blue"))
legend("topright", 
       col = c("green", "blue"),
       lty = 1,
       legend = ssex.colors)



```

```{r}
### PCA reducation chosing axes that explain 80% of the total variance and eigenvalues have to be above 1
#Scree plot shows a lot of the variation is explained in PC1
screeplot(scaled_pca)
abline(a=1,b=0)

## How do we know how many to chose?
summary(scaled_pca)

#Expressed as cummalitve percentages
round(cumsum(100*scaled_pca$CA$eig/sum(scaled_pca$CA$eig)),2)

```

##Practice if we want to pull out the first 3 eigenvalues 
I am not sure how to interpret this because it runs each eigenvalue through but how do we contribute this to a certain trait?
```{r}
data.lm <- lm(scaled_pca$CA$u[,1:3]~ trait$Sex)
summary(data.lm)
```
First three PC account for 88% of the total variation. I would absolutely keep the first three.

#Univariate

Somehow still need to parse the variance for the random effect of family into two groups for each sex and compare that to the model where family is left together. Would nest that give me the same effect?
```{r}
#Filtering traits into male and female lists
str(trait)
trait$Family <- as.factor(trait$Family)
trait$Sex <- as.factor(trait$Sex)

trait %>% 
  filter(Sex == "m") -> male_trait

trait %>% 
  filter(Sex == "f") -> female_trait
```


#Week 3 perimeter
```{r}
#Are these traits sexually dimorphic for their means?
# -6.196064 No affect of sex
Among_Family_P_w_sex <- lmer(Week_3_perimeter~ Sex + (1|Family) + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_P_n_sex <- lmer(Week_3_perimeter~ (1|Family)  + (1|Family:Sex)+ (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

VarCorr(Among_Family_P_n_sex)

summary(Among_Family_P_w_sex)
summary(Among_Family_P_n_sex)

#Comparing models where sex could have an affect. -4.694448
BIC_compare_P_sex_compare <- BIC(Among_Family_P_n_sex)-BIC(Among_Family_P_w_sex)

#Least Square Meanas
output_1_perim <- lsmeans(Among_Family_P_w_sex, ~Sex)

```

```{r}
#Are these traits dimorphic for their variances? 121.3226

trait_1 <- transform(trait,
          obs=factor(1:nrow(trait)),
          Sexdummy=as.numeric(Sex=="m"))
Among_family_P_sex_constrained <- lmer(Week_3_perimeter~Sex+
     (1|Family) + (1|Family:Sex) + (Sexdummy-1|obs), 
     trait_1, REML=TRUE,
     control=lmerControl(check.nobs.vs.nRE  = "ignore",
                         check.nobs.vs.nlev = "ignore")) #constrained variances in lmer

summary(Among_family_P_sex_constrained)

#Compare effect of sex when variance is constrained to be equal

BIC_compare_P_sex_compare_constrained <- BIC(Among_family_P_sex_constrained)-BIC(Among_Family_P_w_sex)

#Least square means and variances
output_2_perim <- lsmeans(Among_family_P_sex_constrained, ~Sex)
```

```{r}
#Does the among family variance differ? -6.348371
Among_Family_P_Fam <- lmer(Week_3_perimeter~ Sex + (1|Family)+ (1|Family:Sex)  + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_P_Fam)

Among_Family_P_n_Fam <- lmer(Week_3_perimeter~ Sex + (1|Family:Sex)  + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

BIC_compare_P_Fam_compare <- BIC(Among_Family_P_n_Fam)-BIC(Among_Family_P_Fam)

VarCorr(Among_Family_P_Fam)


```
```{r}
#Comparing male and female families

#Females
Female_Family_P_Fam <- lmer(Week_3_perimeter~ (1|Family)  + (1|Plate), data = female_trait, REML = FALSE, na.action = na.omit)

Female_Family_P_n_Fam <- lmer(Week_3_perimeter~ (1|Plate), data = female_trait, REML = FALSE, na.action = na.omit)

summary(Female_Family_P_Fam)
summary(Female_Family_P_n_Fam)

BIC_compare_P_female_fam <- BIC(Female_Family_P_n_Fam)-BIC(Female_Family_P_Fam)


#Males
male_Family_P_Fam <- lmer(Week_3_perimeter~ (1|Family)  + (1|Plate), data = male_trait, REML = FALSE, na.action = na.omit)

male_Family_P_n_Fam <- lmer(Week_3_perimeter~ (1|Plate), data = male_trait, REML = FALSE, na.action = na.omit)

summary(male_Family_P_Fam)
summary(male_Family_P_n_Fam)

BIC_compare_P_male_fam <- BIC(male_Family_P_n_Fam)-BIC(male_Family_P_Fam)

```





```{r}
#Does the among family variance for each sex differ? 30.8637
Among_Family_P_w_sex_int <- lmer(Week_3_perimeter~ Sex + (1|Family) + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_P_w_sex_n_int <- lmer(Week_3_perimeter~ Sex +(1|Family) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_P_w_sex_int)
summary(Among_Family_P_w_sex_n_int)


BIC_compare_P_sex_compare_int <- BIC(Among_Family_P_w_sex_n_int)-BIC(Among_Family_P_w_sex_int) #compare


```

```{r}
#Plot the interaction of family and sex.
 means <- 
  trait%>%
  #filter out females 
  filter(Sex=='m')%>%
  group_by(Family)%>%
  summarise(mean=mean(Week_3_perimeter))

myorder_perm <- means$Family[order(means$mean)]

  ggplot(trait, mapping= aes(x =Family,  y = Week_3_perimeter, fill = Sex)) +
        geom_boxplot(alpha=0.7) +
        scale_y_continuous(name = "Perimeter (mm)",
                           breaks = seq(0, 5, 15),
                           limits=c(0, 15)) +
        scale_x_discrete(name = "Month") +
        ggtitle("Boxplot of perimeter per sex") +
        theme_bw() +
        theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
              text = element_text(size = 12, family = "Tahoma"),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(size = 11)) +
        scale_fill_brewer(palette = "Accent")+
        scale_x_discrete(limits=myorder_perm)


```


#Attempting to fix uneven male and female samples.
```{r}


#Boot strapping for unequal numbers of males and females
Family_2= c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29
            ,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45)

#v_family <- array() #empty array for families
v_males<-array() #empty array for males
v_females<-array() #empty array for females
mean_males<-array() #empty array for mean males
mean_females<-array() #empty array for mean females

data_new<-data.frame(matrix(0, ncol = 9, nrow = 1))  #blank matrix to cbind MA 
colnames(data_new) <- colnames(female_trait) #give blank martix correct col.names

for (k in 1:10000){
  
  data_new_female<-data.frame(matrix(0, ncol = 9, nrow = 1))  #blank matrix to cbind MA 
  colnames(data_new_female) <- colnames(trait) #give blank martix correct col.names
  data_new_female$bootrep <- NA #make a blank column for bootrep
  
  for (i in 1:45){
    rand_fam=sample(Family_2,1)
    temp=subset(female_trait, female_trait$Family==rand_fam)
    temp$bootrep<-i
    data_new_female=rbind(data_new_female,temp)
  }
  
  data_new_female<-data_new_female[-1,] #removes seed line of blank data
  Among_Family_P_w_female <- lmer(Week_3_perimeter~ (1|Family) + (1|Family:Sample) + (1|Plate), data = data_new_female, REML = FALSE, na.action = na.omit)
  mean_females[k]=as.data.frame(coef(summary(Among_Family_P_w_female)))[1,1]#gives mean for females maybe
  v_females[k]=as.data.frame(coef(summary(Among_Family_P_w_female)))[1,2]
  
  
  data_new_male<-data.frame(matrix(0, ncol = 9, nrow = 1))  #blank matrix to cbind MA 
  colnames(data_new_male) <- colnames(male_trait) #give blank martix correct col.names
  data_new_male$bootrep <- NA #make a blank column for bootrep
                                                                    
  for (i in 1:45){
    rand_fam=sample(Family_2,1)
    temp=subset(male_trait, male_trait$Family==rand_fam)
    temp$bootrep<-i
    data_new_male=rbind(data_new_male,temp)
  }
  
  data_new_male<-data_new_male[-1,] #removes seed line of blank data
  Among_Family_P_w_male <- lmer(Week_3_perimeter~ (1|Family) + (1|Family:Sample) + (1|Plate), data = data_new_male, REML = FALSE, na.action = na.omit)
  mean_males[k]=as.data.frame(coef(summary(Among_Family_P_w_male)))[1,1]#gives mean for males maybe
  v_males[k]=as.data.frame(coef(summary(Among_Family_P_w_male)))[1,2]
  print(k)
}

output<-data.frame(v_males,v_females,mean_males,mean_females)
write.csv(output, "Bootstrap_growth_perimeter.csv", row.names = FALSE)

```


#Week 3 area
```{r}
#Are these traits sexually dimorphic for their means? -6.90284

Among_Family_A_w_sex <- lmer(Week_3_area~ Sex + (1|Family) + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_A_n_sex <- lmer(Week_3_area~ (1|Family)  + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_A_w_sex)
summary(Among_Family_A_n_sex)

BIC_compare_A_sex_compare <- BIC(Among_Family_A_n_sex)-BIC(Among_Family_A_w_sex)

output_area <- lsmeans(Among_Family_A_w_sex,~ Sex)

```

```{r}
#Are these traits sexually dimorphic for variances 131.1856
#test for differences in variation between sexes in a mixed model (lme4) for week 3 area 

#a   <- lmer (Week_3_area~ Sex + (1|Family) +  (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

#Running constrained variances in lme instead of lmer
#a <- lme(Week_3_area~ Sex,
     #random=~1|Family +  1|Plate,data=trait, method="REML",
     #weights=varIdent(~Sex))


trait_2 <- transform(trait,
          obs=factor(1:nrow(trait)),
          Sexdummy=as.numeric(Sex=="m"))
Among_family_A_w_sex_constrained <- lmer(Week_3_area~Sex+
     (1|Family) + (1|Family:Sex) + (Sexdummy-1|obs), 
     trait_2, REML=TRUE,
     control=lmerControl(check.nobs.vs.nRE  = "ignore",
                         check.nobs.vs.nlev = "ignore")) #constrained variances in lmer

summary(Among_family_A_w_sex_constrained)

#Compare effect of sex when variance is constrained to be equal
#AIC for constrained sex is larger which means that unconstrained is a better fit and therefore variances may be unequal

BIC_compare_A_sex_compare_constrained <- BIC(Among_family_A_w_sex_constrained)-BIC(Among_Family_A_w_sex)

```

```{r}
#Does the among family variance differ? -6.840116
Among_Family_A_Fam <- lmer(Week_3_area~ Sex + (1|Family:Sex) +(1|Family)  + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_A_n_Fam <- lmer(Week_3_area~ Sex + (1|Family:Sex)  + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

BIC_compare_A_Fam_compare <- BIC(Among_Family_A_n_Fam)-BIC(Among_Family_A_Fam)

```

```{r}
#Comparing male and female families

#Females
Female_Family_A_Fam <- lmer(Week_3_area~ (1|Family)  + (1|Plate), data = female_trait, REML = FALSE, na.action = na.omit)

Female_Family_A_n_Fam <- lmer(Week_3_area~ (1|Plate), data = female_trait, REML = FALSE, na.action = na.omit)

summary(Female_Family_A_Fam)
summary(Female_Family_A_n_Fam)

BIC_compare_A_female_fam <- BIC(Female_Family_A_n_Fam)-BIC(Female_Family_A_Fam)


#Males
male_Family_A_Fam <- lmer(Week_3_area~ (1|Family)  + (1|Plate), data = male_trait, REML = FALSE, na.action = na.omit)

male_Family_A_n_Fam <- lmer(Week_3_area~ (1|Plate), data = male_trait, REML = FALSE, na.action = na.omit)

summary(male_Family_A_Fam)
summary(male_Family_A_n_Fam)

BIC_compare_A_male_fam <- BIC(male_Family_A_n_Fam)-BIC(male_Family_A_Fam)

```

```{r}
#Does the among family variance differ between sexes or can sexual dimorphism evolve? 55.37766


Among_Family_A_w_sex_int <- lmer(Week_3_area~ Sex + (1|Family:Sex) +(1|Family)  + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_A_w_sex_n_int <- lmer(Week_3_area~ Sex +(1|Family) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_A_w_sex_int)
summary(Among_Family_A_w_sex_n_int)

BIC_compare_A_sex_compare_int <- BIC(Among_Family_A_w_sex_n_int)-BIC(Among_Family_A_w_sex)

VarCorr(Among_Family_A_w_sex_int)

```

```{r}
#Plot the interaction of family and sex.
 means <- 
  trait%>%
  #filter out females 
  filter(Sex=='m')%>%
  group_by(Family)%>%
  summarise(mean=mean(Week_3_area))

myorder_area <- means$Family[order(means$mean)]

  ggplot(trait, mapping= aes(x =Family,  y = Week_3_area, fill = Sex)) +
        geom_boxplot(alpha=0.7) +
        scale_y_continuous(name = "Area (mm2)",
                           breaks = seq(0, .5, 1),
                           limits=c(0, 1.5)) +
        scale_x_discrete(name = "Month") +
        ggtitle("Boxplot of area per sex") +
        theme_bw() +
        theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
              text = element_text(size = 12, family = "Tahoma"),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(size = 11)) +
        scale_fill_brewer(palette = "Accent")+
        scale_x_discrete(limits=myorder_area)


```

#Days until Gam 
```{r}
#Is there an affact of Sex?  -6.617722

Among_Family_D_w_sex <- lmer(Days_until_gametophores~ Sex + (1|Family) + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_D_n_sex <- lmer(Days_until_gametophores~ (1|Family) + (1|Family:Sex) +(1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_D_w_sex)
summary(Among_Family_D_n_sex)

BIC_compare_D_sex <- BIC(Among_Family_D_n_sex)-BIC(Among_Family_D_w_sex)

output_days <- lsmeans(Among_Family_D_w_sex, ~Sex)
```

```{r}
#Are these traits sexually dimorphic for variances 9.493998
trait_3 <- transform(trait,
          obs=factor(1:nrow(trait)),
          Sexdummy=as.numeric(Sex=="m"))
Among_family_D_w_sex_constrained <- lmer(Days_until_gametophores~Sex+
     (1|Family) + (1|Family:Sex) + (1|Plate)+(Sexdummy-1|obs), 
     trait_3, REML=TRUE,
     control=lmerControl(check.nobs.vs.nRE  = "ignore",
                         check.nobs.vs.nlev = "ignore")) #constrained variances in lmer

summary(Among_family_D_w_sex_constrained)

#Compare effect of sex when variance is constrained to be equal
#AIC for constrained sex is larger which means that unconstrained is a better fit and therefore variances may be unequal

BIC_compare_D_sex_compare_constrained <- BIC(Among_family_D_w_sex_constrained)-BIC(Among_Family_D_w_sex)
```

```{r}
#Does the among family variance differ? -6.467028
Among_Family_D_Fam <- lmer(Days_until_gametophores~ Sex + (1|Family)  +(1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_D_n_Fam <- lmer(Days_until_gametophores~ Sex + (1|Family:Sex)  + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

BIC_compare_D_Fam_compare <- BIC(Among_Family_D_n_Fam)-BIC(Among_Family_D_Fam)

VarCorr(Among_Family_D_Fam)

```

```{r}
#Comparing male and female families

#Females
Female_Family_D_Fam <- lmer(Days_until_gametophores~ (1|Family)  + (1|Plate), data = female_trait, REML = FALSE, na.action = na.omit)

Female_Family_D_n_Fam <- lmer(Days_until_gametophores~ (1|Plate), data = female_trait, REML = FALSE, na.action = na.omit)

summary(Female_Family_D_Fam)
summary(Female_Family_D_n_Fam)

BIC_compare_D_female_fam <- BIC(Female_Family_D_n_Fam)-BIC(Female_Family_D_Fam)


#Males
male_Family_D_Fam <- lmer(Days_until_gametophores~ (1|Family)  + (1|Plate), data = male_trait, REML = FALSE, na.action = na.omit)

male_Family_D_n_Fam <- lmer(Days_until_gametophores~ (1|Plate), data = male_trait, REML = FALSE, na.action = na.omit)

summary(male_Family_D_Fam)
summary(male_Family_D_n_Fam)

BIC_compare_D_male_fam <- BIC(male_Family_D_n_Fam)-BIC(male_Family_D_Fam)

```

```{r}
#Does the among family variance for each sex differ? 16.46835

Among_Family_D_w_sex_int <- lmer(Days_until_gametophores~ Sex  +(1|Family) + (1|Family:Sample) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_D_w_sex_n_int <- lmer(Days_until_gametophores~ Sex + (1|Family)  + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_D_w_sex_int)
summary(Among_Family_D_w_sex_n_int)

BIC_compare_D_int <- BIC(Among_Family_D_w_sex_n_int)-BIC(Among_Family_D_w_sex_int)

```

```{r}

#Plot the interaction of family and sex.
 means <- 
  trait%>%
  #filter out females 
  filter(Sex=='m')%>%
  group_by(Family)%>%
  summarise(mean=mean(Days_until_gametophores))

myorder_days <- means$Family[order(means$mean)]

  ggplot(trait, mapping= aes(x =Family,  y = Days_until_gametophores, fill = Sex)) +
        geom_boxplot(alpha=0.7) +
        scale_y_continuous(name = "Area (mm2)",
                           breaks = seq(0, 50,100),
                           limits=c(0, 150)) +
        ggtitle("Boxplot of Days until gametophores") +
        theme_bw() +
        theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
              text = element_text(size = 12, family = "Tahoma"),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(size = 11)) +
        scale_fill_brewer(palette = "Accent")+
        scale_x_discrete(limits=myorder_days)


```


#Week 3 Circularity
```{r}
#Is circularity sexually dimorphic?  -3.058375
Among_Family_C_w_sex <- lmer(Week_3_Circularity~ Sex  +(1|Family) + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_C_n_sex <- lmer(Week_3_Circularity~ (1|Family) + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_C_w_sex)
summary(Among_Family_C_n_sex)

BIC_compare_C_Sex_compare <- BIC(Among_Family_C_n_sex)-BIC(Among_Family_C_w_sex)

output_circ <- lsmeans(Among_Family_C_w_sex, ~Sex)
```

```{r}
#Are these traits sexually dimorphic for variances? 21.21545
trait_4 <- transform(trait,
          obs=factor(1:nrow(trait)),
          Sexdummy=as.numeric(Sex=="m"))
Among_family_C_w_sex_constrained <- lmer(Week_3_Circularity~Sex+
     (1|Family) + (1|Family:Sex) +(1|Plate)+ (Sexdummy-1|obs), 
     trait_4, REML=TRUE,
     control=lmerControl(check.nobs.vs.nRE  = "ignore",
                         check.nobs.vs.nlev = "ignore")) #constrained variances in lmer

summary(Among_family_C_w_sex_constrained)

#Compare effect of sex when variance is constrained to be equal
#AIC for constrained sex is larger which means that unconstrained is a better fit and therefore variances may be unequal

BIC_compare_C_sex_compare_constrained <- BIC(Among_family_C_w_sex_constrained)-BIC(Among_Family_C_w_sex)
```


```{r}
#Does the among family variance for each sex differ? -2.864805

Among_Family_C_w_sex_fam <- lmer(Week_3_Circularity~ Sex  +(1|Family) +  (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_C_w_sex_n_fam <- lmer(Week_3_Circularity~ Sex   + (1|Family:Sex)+ (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_C_w_sex_fam)
summary(Among_Family_C_w_sex_n_int)

BIC_compare_D_fam <- BIC(Among_Family_C_w_sex_n_fam)-BIC(Among_Family_C_w_sex_fam)


```

```{r}
#Comparing male and female families

#Females
Female_Family_C_Fam <- lmer(Week_3_Circularity~ (1|Family)  + (1|Plate), data = female_trait, REML = FALSE, na.action = na.omit)

Female_Family_C_n_Fam <- lmer(Week_3_Circularity~ (1|Plate), data = female_trait, REML = FALSE, na.action = na.omit)

summary(Female_Family_C_Fam)
summary(Female_Family_C_n_Fam)

BIC_compare_C_female_fam <- BIC(Female_Family_C_n_Fam)-BIC(Female_Family_C_Fam)


#Males
male_Family_C_Fam <- lmer(Week_3_Circularity~ (1|Family)  + (1|Plate), data = male_trait, REML = FALSE, na.action = na.omit)

male_Family_C_n_Fam <- lmer(Week_3_Circularity~ (1|Plate), data = male_trait, REML = FALSE, na.action = na.omit)

summary(male_Family_C_Fam)
summary(male_Family_C_n_Fam)

BIC_compare_C_male_fam <- BIC(male_Family_C_n_Fam)-BIC(male_Family_C_Fam)

```

```{r} 

#Does the among family variance for each sex differ? 0.8656942
Among_Family_C_w_sex_int <- lmer(Week_3_Circularity~ Sex + (1|Family) + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_C_w_sex_n_int <- lmer(Week_3_Circularity~ Sex +(1|Family)  + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_C_w_sex_int)
summary(Among_Family_C_w_sex_n_int)

BIC_compare_C_Sex_int_compare <- BIC(Among_Family_C_w_sex_n_int)-BIC(Among_Family_C_w_sex_int)



```



```{r}

#Plot the interaction of family and sex.
 means <- 
  trait%>%
  #filter out females 
  filter(Sex=='m')%>%
  group_by(Family)%>%
  summarise(mean=mean(Week_3_Circularity))

myorder_circ <- means$Family[order(means$mean)]

  ggplot(trait, mapping= aes(x =Family,  y = Week_3_Circularity, fill = Sex)) +
        geom_boxplot(alpha=0.7) +
        scale_y_continuous(name = "Area (mm2)",
                           breaks = seq(0, 1,2),
                           limits=c(0, 2)) +
        ggtitle("Boxplot of circularity") +
        theme_bw() +
        theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
              text = element_text(size = 12, family = "Tahoma"),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(size = 11)) +
        scale_fill_brewer(palette = "Accent")+
        scale_x_discrete(limits=myorder_circ)

```




#Total days
```{r}
# sexually dimorphic?  -10.27364
Among_Family_T_w_sex <- lmer(Total_Gametophores~ Sex + (1|Family) + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_T_n_sex <- lmer(Total_Gametophores~ (1|Family)  +  (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_T_w_sex)
summary(Among_Family_T_n_sex)

output_total <- lsmeans(Among_Family_T_w_sex, ~Sex)
BIC_compare_T_Sex_compare <- BIC(Among_Family_T_n_sex)-BIC(Among_Family_T_w_sex)
```

```{r}
#Are these traits sexually dimorphic for variances? 9.434474
trait_5 <- transform(trait,
          obs=factor(1:nrow(trait)),
          Sexdummy=as.numeric(Sex=="m"))
Among_family_T_w_sex_constrained <- lmer(Total_Gametophores~Sex+
     (1|Family) + (1|Family:Sex) + (1|Plate)+(Sexdummy-1|obs), 
     trait_5, REML=TRUE,
     control=lmerControl(check.nobs.vs.nRE  = "ignore",
                         check.nobs.vs.nlev = "ignore")) #constrained variances in lmer

summary(Among_family_T_w_sex_constrained)

#Compare effect of sex when variance is constrained to be equal
#AIC for constrained sex is larger which means that unconstrained is a better fit and therefore variances may be unequal

BIC_compare_T_sex_compare_constrained <- BIC(Among_family_T_w_sex_constrained)-BIC(Among_Family_T_w_sex)
```

```{r}

#Does the among family variance for each sex differ? -4.138287
Among_Family_T_w_sex_int <- lmer(Total_Gametophores~ Sex + (1|Family) + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_T_w_sex_n_int <- lmer(Total_Gametophores~ Sex + (1|Family) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_T_w_sex_int)
summary(Among_Family_T_w_sex_n_int)

BIC_compare_T_Sex_int_compare <- BIC(Among_Family_T_w_sex_n_int)-BIC(Among_Family_T_w_sex_int)
```

```{r}
#Comparing male and female families

#Females
Female_Family_T_Fam <- lmer(Total_Gametophores~ (1|Family)  + (1|Plate), data = female_trait, REML = FALSE, na.action = na.omit)

Female_Family_T_n_Fam <- lmer(Total_Gametophores~ (1|Plate), data = female_trait, REML = FALSE, na.action = na.omit)

summary(Female_Family_T_Fam)
summary(Female_Family_T_n_Fam)

BIC_compare_T_female_fam <- BIC(Female_Family_T_n_Fam)-BIC(Female_Family_T_Fam)


#Males
male_Family_T_Fam <- lmer(Total_Gametophores~ (1|Family)  + (1|Plate), data = male_trait, REML = FALSE, na.action = na.omit)

male_Family_T_n_Fam <- lmer(Total_Gametophores~ (1|Plate), data = male_trait, REML = FALSE, na.action = na.omit)

summary(male_Family_T_Fam)
summary(male_Family_T_n_Fam)

BIC_compare_T_male_fam <- BIC(male_Family_T_n_Fam)-BIC(male_Family_T_Fam)

```

```{r}

#Does the family variance for differ? -5.343458

Among_Family_T_w_sex_Fam <- lmer(Total_Gametophores~ Sex + (1|Family) + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

Among_Family_T_w_sex_n_Fam <- lmer(Total_Gametophores~ Sex + (1|Family:Sex) + (1|Plate), data = trait, REML = FALSE, na.action = na.omit)

summary(Among_Family_T_w_sex_Fam)
summary(Among_Family_T_w_sex_n_Fam)

BIC_compare_T_Sex_Fam <- BIC(Among_Family_T_w_sex_n_Fam)-BIC(Among_Family_T_w_sex_Fam)
```


```{r}

#Plot the interaction of family and sex.
 means <- 
  trait%>%
  #filter out females 
  filter(Sex=='m')%>%
  group_by(Family)%>%
  summarise(mean=mean(Total_Gametophores))

myorder_Total <- means$Family[order(means$mean)]

  ggplot(trait, mapping= aes(x =Family,  y = Total_Gametophores, fill = Sex)) +
        geom_boxplot(alpha=0.7) +
        scale_y_continuous(name = "Area (mm2)",
                           breaks = seq(0, 25,30),
                           limits=c(0, 30)) +
        ggtitle("Boxplot of total gametophores") +
        theme_bw() +
        theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
              text = element_text(size = 12, family = "Tahoma"),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(size = 11)) +
        scale_fill_brewer(palette = "Accent")+
        scale_x_discrete(limits=myorder_Total)


```

