---
title: "Leaf_and_reproduciton_data"
author: "Leslie Kollar"
date: "11/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.width = 14, fig.height = 4)
```


```{r libraries, message=FALSE}
library(tidyverse)
library(vegan)
library(plotly)
library(corrplot)
library(ggpubr)
library(Hmisc)
library(lme4)
library(nlme)
library(lsmeans)
theme_set(theme_bw())
```
## Leaf and Reproduction Data
The leaf and reproductive measurements were all taken on the plants with the volatile data so this data matches up exactly. 

```{r dataload}
master <- read.csv("rawdata/LK_master.csv")
```

```{r}
#Selecting only the data we want
#I am not sure if I grabbed the right sex traits but I think these are most interesting yet still need to be better compared. 

master_less <- master[1:619,]%>%
  select(Fam.., Sample.Name, Sample_Sex, Leaf_Length_Average, Leaf_Average_Area, Leaf_perimeter_average,
         Avg_Male_Buds.Stem, Avg_Arch)  


```

```{r}
# Removing DIV/0
str(master_less)
master_less$Leaf_Average_Area <-  as.numeric(str_replace(master_less$Leaf_Average_Area, "#DIV/0", "NA"))
master_less$Leaf_Length_Average <-  as.numeric(str_replace(master_less$Leaf_Length_Average, "#DIV/0", "NA"))
master_less$Leaf_perimeter_average <-  as.numeric(str_replace(master_less$Leaf_perimeter_average, "#DIV/0", "NA"))

#This will put zeros in even for the NAs in the reprodctive measurements
master_less[is.na(master_less)] <- 0

```


```{r}
## 
master_less <- master_less %>% 
  rename(famid = Fam..,
         sampid = Sample.Name,
         ssex = Sample_Sex) %>% 
  mutate(sampid = str_replace_all(sampid, "\\(.*\\)", ""),
      sampid = str_trim(sampid, side = "both"),
       ssex = str_to_lower(as.character(ssex))) %>% # Sex on lower case only
  mutate_at(vars(starts_with("Leaf_")), as.numeric) %>% # Make them all numeric 
  arrange(sampid, .by_group = TRUE) %>% # Order by family number
  ungroup
```

```{r}

#Scaling reproductive traits. I think these traits are much better to use. The other is percent expression and females are more difficult to find. I think these are way more accurate measurements.

#Male clones included
master_less %>% 
  filter(ssex == "m") %>% 
  select(sampid, ssex, famid, Avg_Male_Buds.Stem)%>% 
  drop_na() %>% 
  #group_by(sampid) %>% 
  #summarise(raw_av = mean(Avg_Male_Buds.Stem)) %>% 
  mutate(reprovar = scale( Avg_Male_Buds.Stem),
         ssex = "m")  -> male_reprovar

#Male clones averaged
master_less %>% 
  filter(ssex == "m") %>% 
  select(sampid, ssex, famid, Avg_Male_Buds.Stem)%>% 
  drop_na() %>% 
  group_by(sampid) %>% 
  summarise(raw_av = mean(Avg_Male_Buds.Stem)) %>% 
  mutate(reprovar = scale(raw_av),
         ssex = "m")  -> male_reprovar_avg

#Female clones included
master_less %>% 
  select(sampid, ssex, famid, Avg_Arch) %>% 
  filter(ssex == "f") %>% 
  drop_na() %>% 
  #group_by(sampid) %>% 
  mutate(reprovar = scale(Avg_Arch),
         ssex = "f") -> fem_reprovar

#Female clones averaged
master_less %>% 
  select(sampid, ssex, famid, Avg_Arch) %>% 
  filter(ssex == "f") %>% 
  drop_na() %>% 
  group_by(sampid) %>% 
  summarise(raw_av = mean(Avg_Arch)) %>% 
  mutate(reprovar = scale(raw_av),
         ssex = "f") -> fem_reprovar_avg


#Combining datasets
reprovar <- bind_rows(male_reprovar, fem_reprovar)
reprovar_avg <- bind_rows(male_reprovar_avg, fem_reprovar_avg)
head(reprovar)

#Did not combine with leaf data because different numbers of samples?

```


```{r}
# Visually checking leaf data
master_less[,4:6] %>% 
  gather(key = "trait", value = "value") %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()

```



```{r}

#Averaging clones for leaf data so that we can put leaf and repro in same data sheet if we ever need to maybe for eigenvale reduction. 
master_less %>% 
  select(famid, sampid, ssex, Leaf_Length_Average, Leaf_Average_Area, Leaf_perimeter_average)%>%
  distinct() %>% 
  full_join(., reprovar) %>% 
  distinct() %>%
  set_names(c("Family", "Sample", "Sex","leaf_length", "leaf_area",
              "leaf_perim", "raw_repro", "repro", "female")) -> L_R_Traits

#Removing female reproductive column
L_R_Traits <- L_R_Traits[,1:8]

#Removing raw_repro column
L_R_Traits$raw_repro <- NULL 

L_R_Traits%>%
  distinct()->L_R_Traits_d

saveRDS(L_R_Traits, "cleandata/reproduction_traits.RDS")

```


```{r}
master_less %>% 
  select(-c(famid, sampid, ssex)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")

reprovar %>% 
  select(-c(sampid,ssex, Avg_Male_Buds.Stem, Avg_Arch)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")

L_R_Traits %>% 
  select(-c(Family,Sample,Sex)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")
```


I am not sure when I should scale here. I dont think I should scale the leaf data and the reprod data together because they are not really comparable? Plus the reproductive trait was already scaled to fit into one column. 

The matrix data will be transformed and scaled together in the same data frame!
```{r}
L_R_Traits[is.na(L_R_Traits)] <- 0
scaled_L_R <- scale(L_R_Traits[,4:7])
```

##Running ordination on growth and development traits
```{r}
#Run PCA
scaled_pca <- rda(scaled_L_R)


# Figure for scaled data
biplot(scaled_pca, display = c("sites", 
                   "species"),
       type = c("text",
                "points"))
ssex.colors <- levels(factor(L_R_Traits$Sex))
ordihull(scaled_pca, group = factor(L_R_Traits$Sex),
         col = c("green","blue"))
legend("topright", 
       col = c("green", "blue"),
       lty = 1,
       legend = ssex.colors)


```

```{r}
### PCA reducation chosing axes that explain 80% of the total variance and eigenvalues have to be above 1
#Scree plot shows a lot of the variation is explained in PC1
screeplot(scaled_pca)
abline(a=1,b=0)

## How do we know how many to chose?
summary(scaled_pca)

#Expressed as cummalitve percentages
round(cumsum(100*scaled_pca$CA$eig/sum(scaled_pca$CA$eig)),2)

```

##Practice if we want to pull out the first 2 eigenvalues 
I am not sure how to interpret this because it runs each eigenvalue through but how do we contribute this to a certain trait?
```{r}
data.lm <- lm(scaled_pca$CA$u[,1:2]~ L_R_Traits$Sex)
summary(data.lm)
```
```{r}
master_less %>% 
  filter(ssex == "m") -> male_master_less

master_less %>% 
  filter(ssex == "f") -> female_master_less

L_R_Traits %>% 
  filter(Sex == "m") -> male_L_R_Traits

L_R_Traits  %>% 
  filter(Sex == "f") -> female_L_R_Traits


```


#Univariate analyses
Interpretation: Smaller the AIC and BIC the better

#Leaf length
```{r}

master_less$famid <- as.factor(master_less$famid)

#Are these traits sexually dimorphic for their means? 9.121407
Among_Family_LL_w_sex <- lmer(Leaf_Length_Average~ ssex + (1|famid) +  (1|famid:ssex) , data = master_less, REML = FALSE, na.action = na.omit)

Among_Family_LL_n_sex <- lmer(Leaf_Length_Average~ (1|famid) +  (1|famid:ssex), data = master_less, REML = FALSE, na.action = na.omit)

VarCorr(Among_Family_LL_n_sex)

summary(Among_Family_LL_w_sex)
summary(Among_Family_LL_n_sex)

#Comparing models where sex could have an affect. -4.694448
BIC_compare_LL_sex_compare <- BIC(Among_Family_LL_n_sex)-BIC(Among_Family_LL_w_sex)
output_Llength <- lsmeans(Among_Family_LL_w_sex, ~ssex)

```
```{r}

#Are these traits sexually dimorphic for their variances? 8.196384

trait_L <- transform(master_less,
          obs=factor(1:nrow(master_less)),
          Sexdummy=as.numeric(ssex=="m"))
Among_family_LL_w_sex_constrained <- lmer(Leaf_Length_Average~ssex+
     (1|famid) + (1|famid:ssex) + (Sexdummy-1|obs), 
     trait_L, REML=TRUE,
     control=lmerControl(check.nobs.vs.nRE  = "ignore",
                         check.nobs.vs.nlev = "ignore")) #constrained variances in lmer

summary(Among_family_LL_w_sex_constrained)

#Compare effect of sex when variance is constrained to be equal

BIC_compare_LL_sex_compare_constrained <- BIC(Among_family_LL_w_sex_constrained)-BIC(Among_Family_LL_w_sex)
```


```{r}
#Does the interaction of family and sex have an effect? -0.1316159

Among_Family_LL_n_sex_int<- lmer(Leaf_Length_Average ~ ssex + (1|famid), data = master_less)


Among_Family_LL_w_sex_int<- lmer(Leaf_Length_Average ~ ssex + (1|famid) + (1|famid:ssex), data = master_less)

summary(Among_Family_LL_w_sex_int)
summary(Among_Family_LL_n_sex_int)

BIC_compare_LL_sex_compare_int <- BIC(Among_Family_LL_n_sex_int)-BIC(Among_Family_LL_w_sex_int)

```
```{r}
#Does the among family variance differ? -6.406537
Among_Family_LL_Fam <- lmer(Leaf_Length_Average~ ssex + (1|famid:ssex) +(1|famid)  , data = master_less, REML = FALSE, na.action = na.omit)

Among_Family_LL_n_Fam <- lmer(Leaf_Length_Average~ ssex + (1|famid:ssex) , data = master_less, REML = FALSE, na.action = na.omit)

BIC_compare_LL_Fam_compare <- BIC(Among_Family_LL_n_Fam)-BIC(Among_Family_LL_Fam)

```

```{r}
#Comparing male and female families

#Females
Female_Family_LL_Fam <- lmer(Leaf_Length_Average~ (1|famid) , data = female_master_less, REML = FALSE, na.action = na.omit)

#Female_Family_LL_n_Fam <- lmer(Leaf_Length_Average~, data = female_master_less, REML = FALSE, na.action = na.omit)

summary(Female_Family_LL_Fam)
#summary(Female_Family_LL_n_Fam)

#BIC_compare_LL_female_fam <- BIC(Female_Family_LL_n_Fam)-BIC(Female_Family_LL_Fam)


#Males
male_Family_LL_Fam <- lmer(Leaf_Length_Average~ (1|famid) , data = male_master_less, REML = FALSE, na.action = na.omit)

#male_Family_LL_n_Fam <- lmer(Leaf_Length_Average~ (1|famid), data = male_master_less, REML = FALSE, na.action = na.omit)

summary(male_Family_LL_Fam)
#summary(male_Family_LL_n_Fam)

#BIC_compare_LL_male_fam <- BIC(male_Family_LL_n_Fam)-BIC(male_Family_LL_Fam)

```


```{r}
#Plot the interaction of family and sex.

means <- 
  master_less%>%
  #filter out females 
  filter(ssex=='m')%>%
  group_by(famid)%>%
  summarise(mean=mean(Leaf_Length_Average))

myorder_length_L <- means$famid[order(means$mean)]

  ggplot(master_less, mapping= aes(x =famid,  y = Leaf_Length_Average, fill = ssex)) +
        geom_boxplot(alpha=0.7) +
        scale_y_continuous(name = "Leaf length",
                           breaks = seq(0, 2.5,4.5), 
                           limits=c(0, 4.5)) +
        ggtitle("Boxplot of leaf length per sex") +
        theme_bw() +
        theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
              text = element_text(size = 12, family = "Tahoma"),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(size = 11)) +
        scale_fill_brewer(palette = "Accent")+
        scale_x_discrete(limits=myorder_length_L)


```

#Leaf area
```{r}

#Among family variances for leaf areaa

#Does sex have an affect on means?
Among_Family_LA_w_sex<- lmer(Leaf_Average_Area ~ ssex + (1|famid) + (1|famid:ssex), data = master_less)

Among_Family_LA_n_sex<- lmer(Leaf_Average_Area ~  (1|famid)+ (1|famid:ssex), data = master_less)

summary(Among_Family_LA_w_sex)
summary(Among_Family_LA_n_sex)

BIC_compare_LA_sex_compare <- BIC(Among_Family_LA_n_sex)-BIC(Among_Family_LA_w_sex)
output_Larea <- lsmeans(Among_Family_LA_w_sex, ~ssex)
```

```{r}
#Are these traits sexually dimorphic for their variances? 8.196384

trait_7 <- transform(master_less,
          obs=factor(1:nrow(master_less)),
          Sexdummy=as.numeric(ssex=="m"))
Among_family_LA_w_sex_constrained <- lmer(Leaf_Average_Area~ssex+
     (1|famid) + (1|famid:ssex) + (Sexdummy-1|obs), 
     trait_7, REML=TRUE,
     control=lmerControl(check.nobs.vs.nRE  = "ignore",
                         check.nobs.vs.nlev = "ignore")) #constrained variances in lmer

summary(Among_family_LA_w_sex_constrained)

#Compare effect of sex when variance is constrained to be equal

BIC_compare_LA_sex_compare_constrained <- BIC(Among_family_LA_w_sex_constrained)-BIC(Among_Family_LA_w_sex)

```


```{r}
#Does additive genetic variance for leaf area differ among sexes?
Among_Family_LA_w_sex_int<- lmer(Leaf_Average_Area ~ ssex + (1|famid) + (1|famid:ssex), data = master_less)

Among_Family_LA_n_sex_int<- lmer(Leaf_Average_Area ~ ssex + (1|famid), data = master_less)


summary(Among_Family_LA_w_sex_int)
summary(Among_Family_LA_n_sex_int)

BIC_compare_LA_int_compare <- BIC(Among_Family_LA_w_sex_int)-BIC(Among_Family_LA_n_sex_int)

```

```{r}
#Does the among family variance differ? -6.406537
Among_Family_LA_Fam <- lmer(Leaf_Average_Area~ ssex + (1|famid:ssex) +(1|famid)  , data = master_less, REML = FALSE, na.action = na.omit)

Among_Family_LA_n_Fam <- lmer(Leaf_Average_Area~ ssex + (1|famid:ssex) , data = master_less, REML = FALSE, na.action = na.omit)

BIC_compare_LA_Fam_compare <- BIC(Among_Family_LA_n_Fam)-BIC(Among_Family_LA_Fam)

```

```{r}
#Comparing male and female families

#Females
Female_Family_LA_Fam <- lmer(Leaf_Average_Area~ (1|famid) , data = female_master_less, REML = FALSE, na.action = na.omit)

#Female_Family_LA_n_Fam <- lmer(Leaf_Average_Area~ (1|famid:ssex), data = female_master_less, REML = FALSE, na.action = na.omit)

summary(Female_Family_LA_Fam)
#summary(Female_Family_LA_n_Fam)

#BIC_compare_LA_female_fam <- BIC(Female_Family_LA_n_Fam)-BIC(Female_Family_LA_Fam)


#Males
male_Family_LA_Fam <- lmer(Leaf_Average_Area~ (1|famid) , data = male_master_less, REML = FALSE, na.action = na.omit)

#male_Family_LA_n_Fam <- lmer(Leaf_Average_Area~ (1|famid), data = male_master_less, REML = FALSE, na.action = na.omit)

summary(male_Family_LA_Fam)
#summary(male_Family_LA_n_Fam)

#BIC_compare_LA_male_fam <- (BIC(male_Family_LA_n_Fam)-BIC(male_Family_LA_Fam))

```

```{r}
#Plot the interaction of family and sex.

means <- 
  master_less%>%
  #filter out females 
  filter(ssex=='m')%>%
  group_by(famid)%>%
  summarise(mean=mean(Leaf_Average_Area))

myorder_area_L <- means$famid[order(means$mean)]

  ggplot(master_less, mapping= aes(x =famid,  y = Leaf_Average_Area, fill = ssex)) +
        geom_boxplot(alpha=0.7) +
        scale_y_continuous(name = "Leaf area",
                           breaks = seq(0, .5,1), 
                           limits=c(0, 1)) +
        ggtitle("Boxplot of leaf area per sex") +
        theme_bw() +
        theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
              text = element_text(size = 12, family = "Tahoma"),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(size = 11)) +
        scale_fill_brewer(palette = "Accent")+
        scale_x_discrete(limits=myorder_area_L)


```

#Leaf perimeter
```{r}
#Does Sex have an affect? 25.70148
Among_Family_LP_w_sex<- lmer(Leaf_perimeter_average ~ ssex + (1|famid) + (1|famid:ssex), data = master_less)
Among_Family_LP_n_sex<- lmer(Leaf_perimeter_average ~  (1|famid)+ (1|famid:ssex), data = master_less)

summary(Among_Family_LP_w_sex)
summary(Among_Family_LP_n_sex)

BIC_compare_LP_sex_compare <- BIC(Among_Family_LP_n_sex)-BIC(Among_Family_LP_w_sex)

output_Lperm <- lsmeans(Among_Family_LP_w_sex, ~ssex)
```

```{r}
#Are these traits sexually dimorphic for their variances?  6.428105

trait_8 <- transform(master_less,
          obs=factor(1:nrow(master_less)),
          Sexdummy=as.numeric(ssex=="m"))
Among_family_LP_w_sex_constrained <- lmer(Leaf_perimeter_average~ssex+
     (1|famid) + (1|famid:ssex) + (Sexdummy-1|obs), 
     trait_8, REML=TRUE,
     control=lmerControl(check.nobs.vs.nRE  = "ignore",
                         check.nobs.vs.nlev = "ignore")) #constrained variances in lmer

summary(Among_family_LP_w_sex_constrained)

#Compare effect of sex when variance is constrained to be equal

BIC_compare_LP_sex_compare_constrained <- BIC(Among_family_LP_w_sex_constrained)-BIC(Among_Family_LP_w_sex)
```

```{r}
#Does the among family variance for each sex differ?
Among_Family_LP_w_sex_int<- lmer(Leaf_perimeter_average ~ ssex + (1|famid)  + (1|famid:ssex), data = master_less)

Among_Family_LP_w_sex_n_int<- lmer(Leaf_perimeter_average ~ ssex + (1|famid) , data = master_less)


summary(Among_Family_LP_w_sex_int)
summary(Among_Family_LP_w_sex_n_int)

BIC_compare_LP_sex_compare_int <- BIC(Among_Family_LP_w_sex_n_int)-BIC(Among_Family_LP_w_sex_int)
```


```{r}
#Comparing male and female families

#Females
Female_Family_LP_Fam <- lmer(Leaf_perimeter_average~ (1|famid) , data = female_master_less, REML = FALSE, na.action = na.omit)

#Female_Family_LP_n_Fam <- lmer(Leaf_perimeter_average~ (1|famid:ssex), data = female_master_less, REML = FALSE, na.action = na.omit)

summary(Female_Family_LP_Fam)
#summary(Female_Family_LP_n_Fam)

#BIC_compare_LP_female_fam <- BIC(Female_Family_LP_n_Fam)-BIC(Female_Family_LP_Fam)


#Males
male_Family_LP_Fam <- lmer(Leaf_perimeter_average~ (1|famid) , data = male_master_less, REML = FALSE, na.action = na.omit)

#male_Family_LP_n_Fam <- lmer(Leaf_perimeter_average~ (1|famid), data = male_master_less, REML = FALSE, na.action = na.omit)

summary(male_Family_LP_Fam)
#summary(male_Family_LP_n_Fam)

#BIC_compare_LP_male_fam <- (BIC$male_Family_LP_n_Fam-BIC$male_Family_LP_Fam)

```

```{r}
#Does the among family variance differ? -6.406537
Among_Family_LP_Fam <- lmer(Leaf_perimeter_average~ ssex + (1|famid:ssex) +(1|famid)  , data = master_less, REML = FALSE, na.action = na.omit)

Among_Family_LP_n_Fam <- lmer(Leaf_perimeter_average~ ssex + (1|famid:ssex) , data = master_less, REML = FALSE, na.action = na.omit)

BIC_compare_LP_Fam_compare <- BIC(Among_Family_LP_n_Fam)-BIC(Among_Family_LP_Fam)

```


```{r}
#Plot the interaction of family and sex.

means <- 
  master_less%>%
  #filter out females 
  filter(ssex=='m')%>%
  group_by(famid)%>%
  summarise(mean=mean(Leaf_perimeter_average))

myorder_perm_L <- means$famid[order(means$mean)]

  ggplot(master_less, mapping= aes(x =famid,  y = Leaf_perimeter_average, fill = ssex)) +
        geom_boxplot(alpha=0.7) +
        scale_y_continuous(name = "Leaf perimeter",
                           breaks = seq(0, 5, 25), 
                           limits=c(0, 15)) +
        ggtitle("Boxplot of leaf perimeter per sex") +
        theme_bw() +
        theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
              text = element_text(size = 12, family = "Tahoma"),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(size = 11)) +
        scale_fill_brewer(palette = "Accent")+
        scale_x_discrete(limits=myorder_perm_L)


```



#Repro
```{r} 


#Does sex have an affect?-9.744196
Among_Family_r_w_sex<- lmer(repro ~ Sex + (1|Family), data = L_R_Traits)
Among_Family_r_n_sex<- lmer(repro ~  (1|Family), data = L_R_Traits)

summary(Among_Family_r_w_sex)
summary(Among_Family_r_n_sex)

BIC_compare_r_sex_compare <- BIC(Among_Family_r_n_sex)-BIC(Among_Family_r_w_sex)
output_repro <- lsmeans(Among_Family_r_w_sex, ~Sex)
```

```{r}
#Are these traits sexually dimorphic for their variances? -12.03937

trait_9 <- transform(L_R_Traits,
          obs=factor(1:nrow(L_R_Traits)),
          Sexdummy=as.numeric(Sex=="m"))
Among_family_r_w_sex_constrained <- lmer(repro~Sex+
     (1|Family) + (1|Family:Sex) + (Sexdummy-1|obs), 
     trait_9, REML=TRUE,
     control=lmerControl(check.nobs.vs.nRE  = "ignore",
                         check.nobs.vs.nlev = "ignore")) #constrained variances in lmer

summary(Among_family_r_w_sex_constrained)

#Compare effect of sex when variance is constrained to be equal

BIC_compare_r_sex_compare_constrained <- BIC(Among_family_r_w_sex_constrained)-BIC(Among_Family_r_w_sex)
```

```{r}
#Does the among family variance for each sex differ? 18.91357
Among_Family_r_w_sex_int<- lmer(repro ~ Sex + (1|Family)  + (1|Family:Sex), data = L_R_Traits)

Among_Family_r_w_sex_n_int<- lmer(repro ~ Sex + (1|Family) , data = L_R_Traits)


summary(Among_Family_r_w_sex_int)
summary(Among_Family_r_w_sex_n_int)

BIC_compare_r_sex_compare_int <- BIC(Among_Family_r_w_sex_n_int)-BIC(Among_Family_r_w_sex_int)

```

```{r}

#Does the among family variance for each sex differ? -6.553841
Among_Family_r_w_sex_fam<- lmer(repro ~ Sex + (1|Family)  + (1|Family:Sex), data = L_R_Traits)

Among_Family_r_w_sex_n_fam<- lmer(repro ~ Sex +  (1|Family:Sex), data = L_R_Traits)

summary(Among_Family_r_w_sex_n_fam)
summary(Among_Family_r_w_sex_fam)

BIC_compare_r_sex_compare_Fam <- BIC(Among_Family_r_w_sex_n_fam)-BIC(Among_Family_r_w_sex_fam)

```

```{r}
#Comparing male and female families

#Females
Female_Family_R_Fam <- lmer(repro~ (1|Family:Sex) +(1|Family) , data = female_L_R_Traits, REML = FALSE, na.action = na.omit)

Female_Family_R_n_Fam <- lmer(repro~ (1|Family:Sex), data = female_L_R_Traits, REML = FALSE, na.action = na.omit)

summary(Female_Family_R_Fam)
summary(Female_Family_R_n_Fam)

BIC_compare_R_female_fam <- BIC(Female_Family_R_n_Fam)-BIC(Female_Family_R_Fam)


#Males
male_Family_R_Fam <- lmer(repro~ (1|Family:Sex) +(1|Family), data = male_L_R_Traits, REML = FALSE, na.action = na.omit)


male_Family_R_n_Fam <- lmer(repro~ (1|Family:Sex), data = male_L_R_Traits, REML = FALSE, na.action = na.omit)

summary(male_Family_R_Fam)
summary(male_Family_R_n_Fam)

BIC_compare_R_male_fam <- BIC(male_Family_R_n_Fam)-BIC(male_Family_R_Fam)

```


```{r}
#Plot the interaction of family and sex.
str(L_R_Traits) 
L_R_Traits$Family <- as.character(L_R_Traits$Family)

means <- 
  L_R_Traits%>%
  #filter out females 
  filter(Sex=='m')%>%
  group_by(Family)%>%
  summarise(mean=mean(repro))

myorder_repro <- means$Family[order(means$mean)]

  ggplot(L_R_Traits, mapping= aes(x =Family,  y = repro, fill = Sex)) +
        geom_boxplot(alpha=0.7) +
        scale_y_continuous(name = "Reproductive trait",
                           breaks = seq(0, 5, 25)) +
                           #limits=c(0, 15)) +
        ggtitle("Boxplot of reproductive trait per sex") +
        theme_bw() +
        theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
              text = element_text(size = 12, family = "Tahoma"),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(size = 11)) +
        scale_fill_brewer(palette = "Accent")+
        scale_x_discrete(limits=myorder_repro)


```

#Sex chromosome variation
```{r}
male_master_less$famid <- as.integer(male_master_less$famid)
female_master_less$famid <- as.integer(female_master_less$famid)


#Adding a unique ID to each row. 
male_master_less_2 <- rowid_to_column(male_master_less, "ID")
female_master_less_2 <- rowid_to_column(female_master_less, "ID")

male_master_less_2$famid <- as.character(male_master_less_2$famid)

male_L_R_Traits_2 <- rowid_to_column(male_L_R_Traits, "ID")
female_L_R_Traits_2 <- rowid_to_column(female_L_R_Traits, "ID")



```

```{r}
fem_chromo_Llength <- lm(female_master_less_2$Leaf_Length_Average~ famid + famid:ID, data = female_master_less_2)
summary(fem_chromo_Llength)

plot(fem_chromo_Llength)

male_chromo_Llength <- lm(male_master_less_2$Leaf_Length_Average~ famid + famid:ID + ID, data = male_master_less_2)
summary(male_chromo_Llength)
plot(male_chromo_Larea)

fem_chromo_Larea <- glm(female_master_less_2$Leaf_Average_Area~ famid + famid/ID, data = female_master_less_2)
summary(fem_chromo_Larea)
plot(fem_chromo_Larea)

male_chromo_Larea_1 <- aov(male_master_less_2$Leaf_Average_Area ~ famid + famid/sampid, data = male_master_less_2)
male_chromo_Larea_2 <- Anova(lm(male_master_less_2$Leaf_Average_Area ~ famid + famid/sampid, data = male_master_less_2), type=2)

options(contrasts = c("contr.sum", "contr.poly"))
male_chromo_Larea_3 <- lm(male_master_less_2$Leaf_Average_Area ~  famid + famid/ID, data = male_master_less_2)
Anova(male_chromo_Larea_3, type = "III")
Anova(male_chromo_Larea_3, type = "II")

?summary(male_chromo_Larea)
plot(male_chromo_Larea)

fem_chromo_Lperm <- lm(female_master_less_2$Leaf_perimeter_average~ famid + famid:ID + ID, data = female_master_less_2)
summary(fem_chromo_Lperm)
plot(fem_chromo_Lperm)

male_chromo_Lperm <- lm(male_master_less_2$Leaf_perimeter_average~ famid + famid:ID + ID, data = male_master_less_2)
summary(male_chromo_Lperm)
plot(male_chromo_Lperm)

female_chromo_repo <- lm(female_L_R_Traits_2$repro~ Family + Family:ID + ID, data = female_L_R_Traits_2)
summary(female_chromo_repo)
plot(female_chromo_repo) 

male_chromo_repo <- lm(male_L_R_Traits_2$repro~ Family + Family:ID + ID, data = male_L_R_Traits_2)
summary(male_chromo_repo)
plot(male_chromo_repo)

```









