---
title: "Leaf_and_reproduciton_data"
author: "Leslie Kollar"
date: "11/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.width = 14, fig.height = 4)
```


```{r libraries, message=FALSE}
library(tidyverse)
library(vegan)
library(plotly)
library(corrplot)
library(ggpubr)
library(Hmisc)
theme_set(theme_bw())
```
## Leaf and Reproduction Data
The leaf and reproductive measurements were all taken on the plants with the volatile data so this data matches up exactly. 


```{r dataload}
setwd("/Users/lesliekollar/Desktop/Moss_matrix/")
master <- read.csv("rawdata/LK_master.csv")
```

```{r}
#Selecting only the data we want
#I am not sure if I grabbed the right sex traits but I think these are most interesting yet still need to be better compared. 

master_less <- master%>%
  select(Fam.., Sample.Name, Sample_Sex, Leaf_Length_Average, Leaf_Average_Area, Leaf_perimeter_average,
         Avg_Male_Buds.Stem, Avg_Arch)  

#seq.int(nrow(master_less))
str(master_less)

#Filtering out rows with NAs because scale doees not like NA
master_less %>% 
   filter_all(any_vars(is.na(.)))

master_less[is.na(master_less)] <- 0
```

```{r}

## Something isnt working here...

master_less <- master_less[,1:8] %>% 
  rename(famid = `Fam..`,
         sampid = Sample.Name,
         ssex = Sample_Sex) %>% 
  drop_na(famid) %>% 
  mutate(sampid = str_replace_all(sampid, "\\(.*\\)", ""),
         sampid = str_trim(sampid, side = "both"),
         ssex = str_to_lower(as.character(ssex))) %>% # Sex on lower case only
  mutate_at(vars(starts_with("Leaf_")), as.character) %>%
  mutate_at(vars(starts_with("Leaf_")), as.numeric) %>% # Make them all numeric 
  arrange(sampid, .by_group = TRUE) %>% # Order by family number
  ungroup
```

```{r}

#Scaling reproductive traits. I think these traits are much better to use. The other is percent expression and females are more difficult to find. I think these are way more accurate measurements.

master_less %>% 
  filter(ssex == "m") %>% 
  select(sampid, ssex, Avg_Male_Buds.Stem)%>% 
  drop_na() %>% 
  #group_by(sampid) %>% 
  #summarise(raw_av = mean(Avg_Male_Buds.Stem)) %>% 
  mutate(reprovar = scale( Avg_Male_Buds.Stem),
         ssex = "m")  -> male_reprovar

master_less %>% 
  select(sampid, ssex, Avg_Arch) %>% 
  filter(ssex == "f") %>% 
  drop_na() %>% 
  #group_by(sampid) %>% 
  mutate(reprovar = scale(Avg_Arch),
         ssex = "f") -> fem_reprovar


reprovar <- bind_rows(male_reprovar, fem_reprovar)
head(reprovar)

#Did not combine with leaf data because different numbers of samples?

```
```{r}
# Removing DIV/0
str(master_less)
master_less$Leaf_Average_Area <-  as.numeric(str_replace(master_less$Leaf_Average_Area, "#DIV/0", "NA"))
master_less$Leaf_Length_Average <-  as.numeric(str_replace(master_less$Leaf_Length_Average, "#DIV/0", "NA"))
master_less$Leaf_perimeter_average <-  as.numeric(str_replace(master_less$Leaf_perimeter_average, "#DIV/0", "NA"))



```


```{r}
# Visually checking leaf data
master_less[,4:6] %>% 
  gather(key = "trait", value = "value") %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()

```



```{r}
master_less %>% 
  select(famid, sampid, ssex, Leaf_Length_Average, Leaf_Average_Area, Leaf_perimeter_average) %>% 
  distinct() %>% 
  full_join(., reprovar) %>% 
  distinct() %>%
  set_names(c("Family", "Sample", "Sex","leaf_length", "leaf_area",
              "leaf_perim", "raw_repro", "repro", "female")) -> L_R_Traits

#Removing female reproductive column
L_R_Traits <- L_R_Traits[,1:8]

#Removing raw_repro column
L_R_Traits$raw_repro <- NULL 

saveRDS(L_R_Traits, "cleandata/Final_leaf_reproduction_traits.RDS")

```


```{r}
master_less %>% 
  select(-c(famid, sampid, ssex)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")

reprovar %>% 
  select(-c(sampid,ssex, Avg_Male_Buds.Stem, Avg_Arch)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")

L_R_Traits %>% 
  select(-c(Family,Sample,Sex)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")
```


I am not sure when I should scale here. I dont think I should scale the leaf data and the reprod data together because they are not really comparable? Plus the reproductive trait was already scaled to fit into one column. 

The matrix data will be transformed and scaled together in the same data frame!
```{r}
scaled_L_R <- scale(L_R_Traits[,4:7])
```

##Running ordination on growth and development traits
```{r}
#Run PCA
scaled_pca <- rda(scaled_L_R)


# Figure for log transformed data
biplot(scaled_pca, display = c("sites", 
                   "species"),
       type = c("text",
                "points"))
ssex.colors <- levels(factor(L_R_Traits$Sex))
ordihull(scaled_pca, group = factor(L_R_Traits$Sex),
         col = c("green","blue"))
legend("topright", 
       col = c("green", "blue"),
       lty = 1,
       legend = ssex.colors)


```

```{r}
### PCA reducation chosing axes that explain 80% of the total variance and eigenvalues have to be above 1
#Scree plot shows a lot of the variation is explained in PC1
screeplot(scaled_pca)
abline(a=1,b=0)

## How do we know how many to chose?
summary(scaled_pca)

#Expressed as cummalitve percentages
round(cumsum(100*scaled_pca$CA$eig/sum(scaled_pca$CA$eig)),2)

```

##Practice if we want to pull out the first 2 eigenvalues 
I am not sure how to interpret this because it runs each eigenvalue through but how do we contribute this to a certain trait?
```{r}
data.lm <- lm(scaled_pca$CA$u[,1:2]~ L_R_Traits$Sex)
summary(data.lm)
```














