---
title: "Monster master: we are doing everything here"
output: 
  github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.width = 14, fig.height = 6)

```

```{r libraries}
library(tidyverse)
library(vegan)
library(plotly)
library(corrplot)
library(ggpubr)
library(ggmosaic)

theme_set(theme_bw())
```


# Introduction
This document will contain pretty much everything we have done so far. This works as a draft reference of where we are and what we have done. 

# Data cleaning
Data cleaning is a really important section since lots of our functions will throw errors with NAs, zeroes or any typos. 
```{r}
# Read in the raw data
rawdata <- read.csv("rawdata/LK_master.csv", stringsAsFactors = FALSE)
```


## Trait data
To make things a little easier we are going to tackle the trait data first and then we will go on to clean VOCs. Some changes that we wish to accomplish here include changing the variable names to simpler/shorter ones without special characters. We would also like to remove empty rows at the end, manage the zeroes and NAs. There are some errors in measurements and typos that we need to address. 

```{r}

trait_raw <- rawdata[,1:19] %>% 
  rename(famid = `Fam..`,
         sampid = Sample.Name,
         ssex = Sample_Sex) %>% 
  drop_na(famid) %>% 
  mutate(sampid = str_replace_all(sampid, "\\(.*\\)", ""),
         sampid = str_trim(sampid, side = "both"),
         ssex = str_to_lower(as.character(ssex))) %>% 
  mutate_at(vars(starts_with("Leaf")), as.numeric) %>% 
  group_by(famid) %>% 
  arrange(sampid, .by_group = TRUE) %>% 
  ungroup

```

Fixing some typos or data entry errors:
```{r}
# Fix a data entry error:
# Check Github issue #5
trait_raw$Avg_21_days[which(trait_raw$Avg_21_days == 33)] <- 0.33

# Fix sex data entry error
# Check Github issue #5
trait_raw[which(trait_raw$sampid == "P_18_1_6_B"),]$ssex <- "m"
trait_raw <- trait_raw[-which(trait_raw$sampid == "P_6_6_20"),]
```

#### Leaf traits

```{r}
# Leaf data averaging -----------------------------------------------------

trait_raw[,17:19] %>% 
  gather(key = "trait", value = "value") %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```

There are no clear outliers or errors, so these look ok to average and we create a new data frame with the leaf data. 
```{r}

trait_raw[,c(2, 17:19),] %>% 
  group_by(sampid) %>% 
  summarise_at(vars(starts_with("Leaf")), mean, na.rm = TRUE) -> leaf_data
```

#### Growth and development traits

```{r}
# Growth and Development traits -------------------------------------------

trait_raw[,10:16] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```

There are some crazy outliers here...
```{r}
outliers <- which(trait_raw$Avg_Area._Week_3>50)
trait_raw[outliers,]
```

The error is in the decimal point for the area and circularity: 
```{r}
trait_raw$Avg_Area._Week_3[outliers] <- trait_raw$Avg_Area._Week_3[outliers] / 1000
trait_raw$Avg_Circularity_Week.3[outliers] <- trait_raw$Avg_Circularity_Week.3[outliers] / 1000

# check again 
trait_raw[,10:16] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```

Looks better but, why are there negative values? seems like a typo
```{r}
negs <- which(trait_raw$Avg_Perimeter_Rate < 0)
trait_raw[negs,]

trait_raw$Avg_Perimeter_Rate[negs]
trait_raw$Avg_Perimeter_Rate[negs] <- trait_raw$Avg_Perimeter_Rate[negs] /-10

# Check visually
trait_raw[,10:16] %>% 
  gather(key = "trait", value = "value") %>% 
  drop_na() %>% 
  ggplot(aes(x = trait, y = value)) + 
  geom_boxplot()
```


Now it's ok to get them as family averages:
```{r}
trait_raw[,c(2, 10:16)] %>% 
  drop_na() %>% 
  distinct()-> gro_dev_data
```


#### Reproduction variable
We have different reproduction variables for males and females, so we can scale them and that way keep them under one variable only. This could be wrong, but I honestly don't know what else we could do.
```{r}
# Reproduction variable ---------------------------------------------------

trait_raw %>% 
  filter(ssex == "m") %>% 
  select(sampid, ssex, Avg_Male_Buds.Stem)%>% 
  drop_na() %>% 
  group_by(sampid) %>% 
  summarise(raw_av = mean(Avg_Male_Buds.Stem)) %>% 
  mutate(reprovar = scale(raw_av),
         ssex = "m")  -> male_reprovar

trait_raw %>% 
  select(sampid, ssex, Avg_Arch) %>% 
  filter(ssex == "f") %>% 
  drop_na() %>% 
  group_by(sampid) %>% 
  summarise(raw_av = mean(Avg_Arch)) %>% 
  mutate(reprovar = scale(raw_av),
         ssex = "f") -> fem_reprovar


reprovar <- bind_rows(male_reprovar, fem_reprovar)
reprovar %>% 
  ggplot(aes(x = reprovar, fill = ssex)) +
  geom_histogram()
```

### Clean trait data

We can now join all the smaller datasets and have a master traits data frame. These are not scaled so far and we can evaluate the changes of scaling them vs keeping them as they are.
```{r}
# Join trait data ---------------------------------------------------------

trait_raw %>% 
  select(famid, sampid, ssex) %>% 
  distinct() %>% 
  full_join(., reprovar) %>% 
  full_join(., gro_dev_data) %>%
  full_join(., leaf_data) %>% 
  set_names(c("famid", "sampid", "ssex", "raw_repro", "repro", "area_wk3",
              "perim_wk3", "circ_wk3", "perim_rate", "area_rate",
              "days21", "days_gam", "leaf_length", "leaf_area",
              "leaf_perim")) -> traits
```

#### Histograms raw trait data

```{r}
traits %>% 
  select(-c(famid, sampid, ssex, raw_repro)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#8e9998") +
  facet_wrap( ~ trait, scales = "free")
```

#### Histograms transformed trait data
If we log transform the data and then scale it, the histograms look like this:

```{r}
names(traits)
traits %>% 
  mutate_at(c(5:15), list(~ log10(. + 1))) %>% 
  mutate_at(c(5:15), scale) -> scaled_traits

scaled_traits %>% 
  select(-c(famid, sampid, ssex, raw_repro)) %>% 
  gather(key = "trait", value = "raw_values") %>% 
  ggplot(aes(x = raw_values, y = ..density..)) +
  geom_histogram(na.rm = TRUE, fill = "#7e9bcc") +
  facet_wrap( ~ trait, scales = "free")
```


## Volatile Organic Compounds

Using the same raw data csv file, we are going to select only the identifiers and the VOC data. 

```{r}
# Just keep volatile data and identifiers
# Also, remove the NAs, there's a lot at the end
# Remove '(2)' notation from sample ids

#names(rawdata)
voc_raw <- rawdata[,c(1,2,4,5, 20:109)] %>% 
  rename(famid = `Fam..`,
         sampid = Sample.Name,
         ssex = Sample_Sex) %>% 
  drop_na() %>% 
  mutate(sampid = str_replace_all(sampid, "\\(.*\\)", ""),
         sampid = str_trim(sampid, side = "both"))

#' Shorten voc names
oldnames <- names(voc_raw)
names(voc_raw)[5:94] <- stringr::str_trunc(oldnames[5:94], width = 6, side = "right", ellipsis = "")

```

There was an issue with the notation in one of the compounds, so we are checking that now:
```{r}
#str(voc_raw)
#' m111.0 is a character
#voc_raw$m111.0
#' The notation is different, using a capital E
voc_raw$m111.0 <- as.numeric(str_replace(voc_raw$m111.0, "E", "e"))
```

### May 30 samples
Leslie mentioned that data associated to May 30 was really off, so we are going to check that and probably remove it if the values are not similar to the rest of the dataset. With the figure it is obvious that we need to remove the May 30 data. With these boxplots we see that there are some negative values. These come from subtracting the noise, so any negative values should be zero.

```{r}
voc_raw %>% 
  dplyr::select(-c(famid, sampid, ssex)) %>% 
  gather(., key = "compound", value = "value", -Date.PTR) %>% 
  ggplot(aes(x = Date.PTR, y = value)) +
  geom_boxplot() -> withMay30

voc_raw %>% 
  filter(Date.PTR != "30-May") %>%
  mutate_if(is.numeric, list(~ ifelse(. < 0, 0, .))) %>% 
  arrange(sampid, .by_group = TRUE) %>% 
  dplyr::select(-c(famid, sampid, ssex)) %>% 
  gather(., key = "compound", value = "value", -Date.PTR) %>% 
  ggplot(aes(x = Date.PTR, y = value)) +
  geom_boxplot() ->woutMay30

ggarrange(withMay30, woutMay30, nrow = 2)

```

### Using a 10% threshold
We are only going to keep the compounds for which at least 10% of the samples have a value greater than zero. 

```{r}
voc_raw %>% 
  filter(Date.PTR != "30-May") %>%
  mutate_if(is.numeric, list(~ ifelse(. < 0, 0, .))) %>% 
  arrange(sampid, .by_group = TRUE) %>% 
  dplyr::select(-Date.PTR) -> vocs

# Create a dataset with only the compounds for which at least 10% of the observations show it.
vocs %>% 
  select(-famid, -sampid, -ssex) %>% 
  mutate_if(is.numeric, function(x) ifelse(x > 0, 1, 0)) %>% 
  colSums() %>% 
  data.frame() %>% 
  rownames_to_column() %>%
  setNames(., c("voc", "count")) %>% 
  mutate(prcnt = count/nrow(vocs)) %>% 
  filter(prcnt >= 0.1) -> voc_filter

vocs %>% 
  select(famid, sampid, ssex, c(voc_filter$voc)) %>% 
  group_by(sampid) %>%
  mutate_at(vars(starts_with("m")), max) %>% 
  ungroup () %>% 
  distinct() %>% 
  mutate(ssex = str_to_lower(as.character(ssex))) -> voc_data

```

### VOC Clustering

For us to cluster the VOC data, we need to look at it first. 

```{r}
long_voc_data <- voc_data %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))

```

#### VOC data exploration
We know we are working with very small numbers and this influences how we will manage the data and what sort of clustering we can do for it. With a quick check, we see that the numbers are very small and that different compounds have very different ranges as well.
```{r}
long_voc_data %>% 
  ggplot(aes(x = voc, y = conc)) + 
  geom_point() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90))
```

After looking at the data, the first thought is to log transform it. To check for any measurement outliers we might want to get rid of the zeroes and transform the data to identify potential outliers.

```{r}
long_voc_data %>% 
  filter(conc > 0) %>%
  ggplot(aes(x = voc, y = conc)) + 
  geom_boxplot() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10()
```

With the information, we can clean the data again and remove these outliers.
```{r}
# Making these outliers be zero
voc_data[voc_data < 10e-15] <- 0

long_voc_data <- voc_data %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))


```

```{r eval = FALSE}
long_voc_data %>% 
  ggplot(aes(x = voc, y = conc)) + 
  geom_boxplot() +
  theme_bw() + 
  lims(y = c(1e-13, 1e-05)) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10()
```
#### Transformation and scaling

Based on PTR machines detection threshold, any values smaller than `1e-12` are equivalent to zero. Then we will standardize the data.
```{r}
log_vocs <- voc_data %>% 
  mutate_at(vars(starts_with("m")), list(~ log10(. + 1e-12)))

st_log_vocs <- log_vocs %>% 
  mutate_at(vars(starts_with("m")), scale)

long_vocs <- st_log_vocs %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))

```

This is how the transformed data looks like now:

```{r fig.width=24}

long_vocs %>% 
  ggplot(aes(x = voc, y = conc)) +
  geom_boxplot(notch = TRUE, alpha = 0.2) +
  labs(x = "VOC ID",
       y = "Log transformed and scaled concentration") +
  theme(axis.text.x = element_text(angle = 90))

long_vocs %>% 
  ggplot(aes(x = voc, y = conc, fill = ssex)) +
  geom_boxplot(notch = TRUE, alpha = 0.2) +
  labs(x = "VOC ID",
       y = "Log transformed and scaled concentration") +
  theme(axis.text.x = element_text(angle = 90))
  
```

#### Correlations

```{r}
all_corr <- cor(st_log_vocs[,4:78])
corrplot(all_corr, method = "circle", type = "upper", tl.col = "black", tl.srt = 45, tl.cex = .3, order = "hclust", mar=c(0,0,2,0))
```

#### Clustering

Number of clusters is kind of arbitrary

```{r}
voc_dist <- hclust(dist(all_corr))
plot(voc_dist)
rect.hclust(voc_dist, k =15)
```

Using the clustering to create a new data frame:

```{r}

voc_clusters <- data.frame(clust = cutree(voc_dist, k = 15)) %>% 
  rownames_to_column(var = "voc")

long_vocs %>% 
  left_join(voc_clusters) %>% 
  group_by(sampid, clust, famid, ssex) %>% 
  summarise(voc_value = mean(conc)) %>% 
  mutate(voc_clust = str_pad(clust, width = 2, side = "left", pad = 0),
         voc_clust = paste0("clust_", voc_clust)) %>% 
  ungroup() %>% 
  select(-clust) -> clustered_long_vocs

clustered_long_vocs %>% 
  spread(key = voc_clust, value = voc_value) -> clustered_voc_data
```

```{r}
clustered_long_vocs %>% 
  ggplot(aes(x = voc_clust, y = voc_value, fill = ssex)) +
  geom_boxplot(notch = TRUE, alpha = 0.2) +
  labs(x = "VOC ID",
       y = "Log transformed and scaled concentration") +
  theme(axis.text.x = element_text(angle = 90))
```



