---
title: "Volatile organic compounds clustering/ordination"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

```

```{r libraries}
library(tidyverse)
library(vegan)
library(plotly)
library(corrplot)
```


## Data transformation
The clean RDS data file should be available locally after you run the `master_cleanup.R` file.This will take the master csv data file and output the RDS we use here. The data we will be using has no duplicates and uses a 10% threshold. This threshold establishes the minimum number of samples that would show a given compound for that compound to be included in the analysis.

```{r dataload}
# source("R/master_cleanup.R")
voc_data <- readRDS("cleandata/voc_10.RDS")
```

```{r dataVis}
# gather the data
# do a bloxplot for each compound in the x axis.
# Would show very small values and a large number of zeroes.
# Maybe just use the positive values, and show number of obs at the top. 
# this from the seed dispersal paper.

```


I am thinking that we can log transform the data, and add a really small value, a couple of orders of magnitude from the minimum value in the data set. Once it is in a log scale we can shift the values so that they are not negative but are on the positive side of the spectrum.















