---
title: "Volatile organic compounds clustering/ordination"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 14, fig.height = 6)

```

```{r libraries}
library(tidyverse)
library(vegan)
library(plotly)
library(corrplot)
library(ggpubr)
library(ggmosaic)
```

The clean RDS data file should be available locally after you run the `master_cleanup.R` file.This will take the master csv data file and output the RDS we use here. The data we will be using has no duplicates and uses a 10% threshold. This threshold establishes the minimum number of samples that would show a given compound for that compound to be included in the analysis.

```{r dataload}
# source("R/master_cleanup.R")
voc_data <- readRDS("cleandata/voc_clean10.RDS")

# this is to use ggplot
long_voc_data <- voc_data %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))

```

## Data exploration
We know we are working with very small numbers and this influences how we will manage the data and what sort of clustering we can do for it. With a quick check, we see that the numbers are very small and that different compounds have very different ranges as well.  
```{r}
long_voc_data %>% 
  ggplot(aes(x = voc, y = conc)) + 
  geom_point() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90))
```

We also know that there are a significant number of zeros in the data, and if we visualize it without the zeroes, it can give us some insight regarding the abundant compounds which we are actually interested in. However, when you compare these two figures, they don't seem too different

```{r}
long_voc_data %>% 
  filter(conc > 0) %>% 
  ggplot(aes(x = voc, y = conc)) + 
  geom_point() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90))

```

We might want to get a hint of what the data looks like for males and females and how these might differ by family.  

```{r}

long_voc_data %>% 
  filter(conc > 0) %>% 
  mutate(famid = as.numeric(famid)) %>% 
  ggplot(., aes(x = voc, y = conc, color = famid)) +
  geom_point(alpha = 0.4) +
  scale_color_viridis_c() +
  facet_wrap(~ ssex) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

```


```{r facetPlot, fig.height=12}
long_voc_data %>% 
  filter(conc > 0) %>% 
  ggplot(aes(x = voc, y = conc, color = ssex)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~famid, ncol = 6) +
  theme(axis.text.x = element_blank())
```

## Data transformation

With these figures it is very clear that the compounds are at different scales, so we might want to log transform them and check the result, to compare with the first figure. I have chosen not to include the zeroes here since log10(0) = -inf. This actually makes it easy to see that there are some outliers that should probably just be considered a zero. It is pretty clear that most of the compounds we are interested in are in a range.


```{r}
long_voc_data %>% 
  filter(conc > 0) %>%
  ggplot(aes(x = voc, y = conc)) + 
  geom_boxplot() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10()
```



With the information, we can clean the data again and remove these outliers.
```{r}
voc_data[voc_data < 10e-15] <- 0
long_voc_data <- voc_data %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))

long_voc_data %>% 
  filter(conc > 0) %>%
  ggplot(aes(x = voc, y = conc)) + 
  geom_point() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10()
```


```{r fig.height=12}
long_voc_data %>% 
  filter(conc > 0) %>% 
  ggplot(aes(x = voc, y = conc, color = ssex)) +
  geom_point(alpha = 0.5) +
  scale_y_log10() +
  facet_wrap(~famid, ncol = 6) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```
Ok, this is good! and now we have a better idea of the data and how we can work with it for any ordination type of analysis.
I am thinking that we can log transform the data, and add a really small value, a couple of orders of magnitude from the minimum value in the data set. Once it is in a log scale we can shift the values so that they are not negative but are on the positive side of the spectrum. This is not taking the absolute value but shifting with a sum so that the biggest values correspond to the largest concentrations of the compounds and smaller values correspond to smaller concentrations. If we were to take the absolute value this concept would be inverse and less intuitive. The following histograms are examples of these transformations using the data for compound **m87.04**

```{r}
long_voc_data %>% 
  filter(voc == "m87.04") -> m87.04

# Raw distribution
m87.04 %>% 
  ggplot() +
  geom_histogram(aes(x = conc)) + 
  labs(title = "Raw distribution") -> raw

# log distribution, there are 3 zeroes that get removed
m87.04 %>% 
  mutate(conc = log10(conc)) %>% 
  ggplot() +
  geom_histogram(aes(x = conc)) + 
  labs(title = "Log transformed - zero removed") -> logh

# if we add the minimum value so we don't loose the zeroes
head(sort(m87.04$conc))
m87.04 %>% 
  mutate(conc = log10(conc + 1e-11)) %>% 
  ggplot() +
  geom_histogram(aes(x = conc)) +
  labs(title = "log plus smallest value") -> logz

# if to the one we added the minimum we shift it to all values are positive
m87.04 %>% 
  mutate(conc = log10(conc + 1e-11) + 11) %>%  
  ggplot() +
  geom_histogram(aes(x = conc)) +
  labs(title = "shifted log plus value") -> logzplus

# Take absolute value is like mirror, so all are positives
m87.04 %>% 
  mutate(conc = abs(log10(conc + 1e-11))) %>%  
  ggplot() +
  geom_histogram(aes(x = conc)) +
  labs(title = "abs log plus value") -> logzabs


ggarrange(raw, logh, logz, logzabs, ncol = 2, nrow = 2)

```




## Ordination
We will start with a PCA, since the goal of that is to reduce the number of dimensions and see what compounds might be acting in a similar way. The hulls show something we were able to see before, and it's that the males and females overlap on their volatile profiles.


```{r}
# Use the transformation of data
voc_data %>%
  select(-c(sampid, famid, ssex)) %>% 
  mutate_all(., list( ~ ifelse(. == 0, 1, .))) %>% 
  min() -> minvoc

zeroes <- 10^(round(log10(minvoc)) - 1)

voc_ordination_data <- voc_data %>% 
  #select(-c(sampid, famid, ssex)) %>% 
  mutate_at(vars(starts_with("m")), list( ~ log10(. + zeroes) - log10(zeroes)))

```

This is how the transformed data looks now.

```{r fig.height=12}
voc_ordination_data %>%
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc))) %>%
  ggplot(aes(x = voc, y = conc, color = ssex)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~famid, ncol = 6) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```

```{r fig.height=7, fig.width=7}
voc_pca <- rda(voc_ordination_data[,4:78])

biplot(voc_pca, display = c("sites", 
                   "species"),
       type = c("text",
                "points"))
ssex.colors <- levels(factor(voc_ordination_data$ssex))
# add hulls
ordihull(voc_pca, group = factor(voc_ordination_data$ssex),
         col = c("blue","black"))
legend("topright", 
       col = c("blue", "black"),
       lty = 1,
       legend = ssex.colors)
```

We can try running a detrended correspondence analysis, but it shows pretty much a similar trend on how the volatiles show up together. It seems to have some influence with their mass, which is kind of intuitive.

```{r fig.height=7, fig.width=7}
dca_voc <- decorana(voc_ordination_data[,4:78])
plot(dca_voc, display = "species")
```


## NMDS
We can try an NMDS following the process Leslie had done in the past, which still doesn't show convergence.

```{r}
voc_nmds <- metaMDS(voc_ordination_data[,4:78])
```

```{r}
scores <- as.data.frame(scores(voc_nmds)) %>% 
  mutate(samplenum = rownames(.),
         family = factor(voc_ordination_data$famid),
         sex = factor(voc_ordination_data$ssex))


vol.scores <- voc_nmds$species %>% 
  as.data.frame() %>% 
  mutate(vol.ident = rownames(.))

ggplot() +
  theme_bw() +
  #geom_text(data = vol.scores, aes(x = MDS1, y = MDS2, label = vol.ident)) +
  geom_point(data = scores, aes(x = NMDS1, y = NMDS2, color = family)) +
  scale_color_viridis_d() +
  #scale_color_viridis_c() +
  coord_equal()
```

The hulls by sex


```{r}

scores %>% 
  group_by(sex) %>%
  nest() %>% 
  mutate(
    hull = map(data, ~ with(., chull(NMDS1, NMDS2))),
    out = map2(data, hull, ~ .x[.y,,drop=FALSE])
  ) %>% 
  select(-data) %>% 
  unnest() -> hullData


ggplot() +
  theme_bw() +
  #geom_text(data = spp_scores, aes(x = MDS1, y = MDS2, label = species)) +
  #geom_text(data = scores, aes(x = NMDS1, y = NMDS2, label = site)) + 
  geom_polygon(data = hullData, aes(x = NMDS1, y = NMDS2, fill = sex, group = sex), alpha = 0.2) +
  geom_point(data = scores, aes(x = NMDS1, y = NMDS2, color = sex)) +
  coord_equal() +
  scale_fill_viridis_d() +
  scale_color_viridis_d()
```

## Correlations

These side by side correlations are donw for all of them, then by sex. The first panel will show the correlations organized by mass, whereas the second panel will use a clustering algorithm. The first set of correlation matrices is using the transformed data.
```{r}
# Correlation using the transformed data

corr_data <- voc_ordination_data

voc_corr <- cor(corr_data[4:78])
par(mfrow = c(1,2))
corrplot(voc_corr, method = "circle", type = "upper", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "All - not clust")
corrplot(voc_corr, method = "circle", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "All - hclust")


x <- corr_data %>% 
  filter(ssex == "m")
voc_corr <- cor(x[4:78])
par(mfrow = c(1,2))
corrplot(voc_corr, method = "circle", type = "upper", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "Males - not clust")
corrplot(voc_corr, method = "circle", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "Males - hclust")

x <- corr_data %>% 
  filter(ssex == "f")
voc_corr <- cor(x[4:78])
par(mfrow = c(1,2))
corrplot(voc_corr, method = "circle", type = "upper", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "Females - not clust")
corrplot(voc_corr, method = "circle", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "Females - hclust")

```

This next set of correlation matrices is using the raw data, without any transformation
```{r}
# Correlation using the transformed data

corr_data <- voc_data

voc_corr <- cor(corr_data[4:78])
par(mfrow = c(1,2))
corrplot(voc_corr, method = "circle", type = "upper", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "All - not clust")
corrplot(voc_corr, method = "circle", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "All - hclust")


x <- corr_data %>% 
  filter(ssex == "m")
voc_corr <- cor(x[4:78])
par(mfrow = c(1,2))
corrplot(voc_corr, method = "circle", type = "upper", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "Males - not clust")
corrplot(voc_corr, method = "circle", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "Males - hclust")

x <- corr_data %>% 
  filter(ssex == "f")
voc_corr <- cor(x[4:78])
par(mfrow = c(1,2))
corrplot(voc_corr, method = "circle", type = "upper", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "Females - not clust")
corrplot(voc_corr, method = "circle", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45, tl.cex = .3, title = "Females - hclust")

```


We can tell there are some compounds that can be grouped, this compares the transformed data and the original raw data
```{r}
par(mfrow = c(1,2))
voc_corr <- cor(voc_ordination_data[,4:78])
corrplot(voc_corr, method = "circle", order = "hclust", tl.col = "black", tl.cex = .5, addrect = 10, title = "Raw all")

voc_corr <- cor(voc_data[,4:78])
corrplot(voc_corr, method = "circle", order = "hclust", tl.col = "black", tl.cex = .5, addrect = 16, title = "Transformed")



```

Comparing these two correlation matrices is equivalent to comparing these two datasets:


```{r}

voc_ordination_data %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc))) %>% 
  ggplot(aes(x = voc, y = conc)) + 
  geom_point() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) -> logplot

long_voc_data %>% 
  #filter(conc > 0) %>%
  ggplot(aes(x = voc, y = conc)) + 
  geom_point() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) -> rawplot

ggarrange(logplot, rawplot, nrow = 2)
```


Enough of visualizing things. So now let's get some clustering going so we can determine what groups can go together. A question with this is the effect of a transformation on the data and the clustering. And with that, also considering whether we want to take a weighted average of the raw data or the transformed data for each cluster. I would suggest taking the weighted average of the raw data, and with this is probably following the cluster analysis on raw data. The reasoning behind this is that if we take a weighted average we would be going around the issue of having zeroes. They would just get mixed into the averaging process. On the other hand, if some clusters have only 2 or 3 compounds, a zero would have a big effect on the output, but then again, that's why we would be doing weighted averages.

After giving it a little more thought, maybe we want to consider the transformed data instead. If you focus on the cluster from the visualization, it seems like the big cluster from the raw data might just be those compounds that have really small concentrations. So then, these clusters are really just based on the concentrations, so obviously the ones that have super low concentrations might group together. By transforming the data is like we have all the compounds on a similar scale and we can actually compare them to each other and have correlations that make sense on a unified scale.

```{r}
voc_dist <- dist(voc_ordination_data[,4:78], method = "euclidean")
clust_fit <- hclust(voc_dist, method = "ward")
plot(clust_fit)

clust_groups <- cutree(clust_fit, k=2)
rect.hclust(clust_fit, k = 2, border = "red")


voc_ordination_data$cluster <- clust_groups

# table(voc_ordination_data$ssex, voc_ordination_data$cluster)
```



It seems like that is clustering the samples, not the volatiles. So maybe the focus is with correlation matrix and then calculate the distance


```{r}
# For the rawdata
# all_corr <- cor(voc_data[,4:78])

# Or the transformed data
all_corr <- cor(voc_ordination_data[,4:78])

all_dis <- dist(all_corr)
all_clust <- hclust(all_dis)
plot(all_clust)
rect.hclust(all_clust, k = 10)


voc_groups <- cbind(cutree(all_clust, k = 10), rownames(all_corr))

```


Now, I need to use the clusters given in the dendrogram to group the volatiles and calculate weighted averages for each sample.







