---
title: "Volatile organic compounds clustering/ordination"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 14, fig.height = 6)

```

```{r libraries}
library(tidyverse)
library(vegan)
library(plotly)
library(corrplot)
library(ggpubr)
```

The clean RDS data file should be available locally after you run the `master_cleanup.R` file.This will take the master csv data file and output the RDS we use here. The data we will be using has no duplicates and uses a 10% threshold. This threshold establishes the minimum number of samples that would show a given compound for that compound to be included in the analysis.

```{r dataload}
# source("R/master_cleanup.R")
voc_data <- readRDS("cleandata/voc_clean10.RDS")

# this is to use ggplot
long_voc_data <- voc_data %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))

```

## Data exploration
We know we are working with very small numbers and this influences how we will manage the data and what sort of clustering we can do for it. With a quick check, we see that the numbers are very small and that different compounds have very different ranges as well.  
```{r}
long_voc_data %>% 
  ggplot(aes(x = voc, y = conc)) + 
  geom_point() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90))
```

We also know that there are a significant number of zeros in the data, and if we visualize it without the zeroes, it can give us some insight regarding the abundant compounds which we are actually interested in. However, when you compare these two figures, they don't seem too different

```{r}
long_voc_data %>% 
  filter(conc > 0) %>% 
  ggplot(aes(x = voc, y = conc)) + 
  geom_point() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90))

```

We might want to get a hint of what the data looks like for males and females and how these might differ by family.  

```{r}

long_voc_data %>% 
  filter(conc > 0) %>% 
  mutate(famid = as.numeric(famid)) %>% 
  ggplot(., aes(x = voc, y = conc, color = famid)) +
  geom_point(alpha = 0.4) +
  scale_color_viridis_c() +
  facet_wrap(~ ssex) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

```


```{r facetPlot, fig.height=12}
long_voc_data %>% 
  filter(conc > 0) %>% 
  ggplot(aes(x = voc, y = conc, color = ssex)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~famid, ncol = 6) +
  theme(axis.text.x = element_blank())
```

## Data transformation

With these figures it is very clear that the compounds are at different scales, so we might want to log transform them and check the result, to compare with the first figure. I have chosen not to include the zeroes here since log10(0) = -inf. This actually makes it easy to see that there are some outliers that should probably just be considered a zero. It is pretty clear that most of the compounds we are interested in are in a range.


```{r}
long_voc_data %>% 
  filter(conc > 0) %>%
  ggplot(aes(x = voc, y = conc)) + 
  geom_boxplot() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10()
```



With the information, we can clean the data again and remove these outliers.
```{r}
voc_data[voc_data < 10e-15] <- 0
long_voc_data <- voc_data %>% 
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc)))

long_voc_data %>% 
  filter(conc > 0) %>%
  ggplot(aes(x = voc, y = conc)) + 
  geom_point() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10()
```


```{r fig.height=12}
long_voc_data %>% 
  filter(conc > 0) %>% 
  ggplot(aes(x = voc, y = conc, color = ssex)) +
  geom_point(alpha = 0.5) +
  scale_y_log10() +
  facet_wrap(~famid, ncol = 6) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```
Ok, this is good! and now we have a better idea of the data and how we can work with it for any ordination type of analysis.
I am thinking that we can log transform the data, and add a really small value, a couple of orders of magnitude from the minimum value in the data set. Once it is in a log scale we can shift the values so that they are not negative but are on the positive side of the spectrum. This is not taking the absolute value but shifting with a sum so that the biggest values correspond to the largest concentrations of the compounds and smaller values correspond to smaller concentrations. If we were to take the absolute value this concept would be inverse and less intuitive. The following histograms are examples of these transformations using the data for compound **m87.04**

```{r}
long_voc_data %>% 
  filter(voc == "m87.04") -> m87.04

# Raw distribution
m87.04 %>% 
  ggplot() +
  geom_histogram(aes(x = conc)) + 
  labs(title = "Raw distribution") -> raw

# log distribution, there are 3 zeroes that get removed
m87.04 %>% 
  mutate(conc = log10(conc)) %>% 
  ggplot() +
  geom_histogram(aes(x = conc)) + 
  labs(title = "Log transformed - zero removed") -> logh

# if we add the minimum value so we don't loose the zeroes
head(sort(m87.04$conc))
m87.04 %>% 
  mutate(conc = log10(conc + 1e-11)) %>% 
  ggplot() +
  geom_histogram(aes(x = conc)) +
  labs(title = "log plus smallest value") -> logz

# if to the one we added the minimum we shift it to all values are positive
m87.04 %>% 
  mutate(conc = log10(conc + 1e-11) + 11) %>%  
  ggplot() +
  geom_histogram(aes(x = conc)) +
  labs(title = "shifted log plus value") -> logzplus

ggarrange(raw, logh, logz, logzplus, ncol = 2, nrow = 2)

```




## Ordination
We will start with a PCA, since the goal of that is to reduce the number of dimensions and see what compounds might be acting in a similar way. The hulls show something we were able to see before, and it's that the males and females overlap on their volatile profiles.


```{r}
# Use the transformation of data
voc_data %>%
  select(-c(sampid, famid, ssex)) %>% 
  mutate_all(., list( ~ ifelse(. == 0, 1, .))) %>% 
  min() -> minvoc

zeroes <- 10^(round(log10(minvoc)) - 1)

voc_ordination_data <- voc_data %>% 
  #select(-c(sampid, famid, ssex)) %>% 
  mutate_at(vars(starts_with("m")), list( ~ log10(. + zeroes) - log10(zeroes)))

```

This is how the transformed data looks now.

```{r fig.height=12}
voc_ordination_data %>%
  gather(., key = "voc", value = "conc", -c(famid, sampid, ssex)) %>% 
  mutate(voc = factor(voc, levels = unique(voc))) %>%
  ggplot(aes(x = voc, y = conc, color = ssex)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~famid, ncol = 6) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```

```{r fig.height=7, fig.width=7}
voc_pca <- rda(voc_ordination_data[,4:78])

biplot(voc_pca, display = c("sites", 
                   "species"),
       type = c("text",
                "points"))
ssex.colors <- levels(factor(voc_ordination_data$ssex))
# add hulls
ordihull(voc_pca, group = factor(voc_ordination_data$ssex),
         col = c("blue","black"))
legend("topright", 
       col = c("blue", "black"),
       lty = 1,
       legend = ssex.colors)
```

We can try running a detrended correspondence analysis, but it shows pretty much a similar trend on how the volatiles show up together. It seems to have some influence with their mass, which is kind of intuitive.

```{r fig.height=7, fig.width=7}
dca_voc <- decorana(voc_ordination_data[,4:78])
plot(dca_voc, display = "species")
```


## NMDS
We can try an NMDS following the process Leslie had done in the past, which still doesn't show convergence.

```{r}
voc_nmds <- metaMDS(voc_ordination_data[,4:78])
```

```{r}
scores <- as.data.frame(scores(voc_nmds)) %>% 
  mutate(samplenum = rownames(.),
         family = factor(voc_ordination_data$famid),
         sex = factor(voc_ordination_data$ssex))


vol.scores <- voc_nmds$species %>% 
  as.data.frame() %>% 
  mutate(vol.ident = rownames(.))

ggplot() +
  theme_bw() +
  #geom_text(data = vol.scores, aes(x = MDS1, y = MDS2, label = vol.ident)) +
  geom_point(data = scores, aes(x = NMDS1, y = NMDS2, color = family)) +
  scale_color_viridis_d() +
  #scale_color_viridis_c() +
  coord_equal()
```

The hulls by sex


```{r}

scores %>% 
  group_by(sex) %>%
  nest() %>% 
  mutate(
    hull = map(data, ~ with(., chull(NMDS1, NMDS2))),
    out = map2(data, hull, ~ .x[.y,,drop=FALSE])
  ) %>% 
  select(-data) %>% 
  unnest() -> hullData


ggplot() +
  theme_bw() +
  #geom_text(data = spp_scores, aes(x = MDS1, y = MDS2, label = species)) +
  #geom_text(data = scores, aes(x = NMDS1, y = NMDS2, label = site)) + 
  geom_polygon(data = hullData, aes(x = NMDS1, y = NMDS2, fill = sex, group = sex), alpha = 0.2) +
  geom_point(data = scores, aes(x = NMDS1, y = NMDS2, color = sex)) +
  coord_equal() +
  scale_fill_viridis_d() +
  scale_color_viridis_d()
```

## Correlations

For all of them
```{r}
# Correlation
x <- voc_ordination_data
voc_corr <- cor(x[4:78])
par(mfrow = c(1,2))
corrplot(voc_corr, method = "circle", type = "upper", tl.col = "black", tl.srt = 45, tl.cex = .3)
corrplot(voc_corr, method = "circle", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45, tl.cex = .3)
```


For males
```{r}
# Correlation
x <- voc_ordination_data %>% 
  filter(ssex == "m")
voc_corr <- cor(x[4:78])
par(mfrow = c(1,2))
corrplot(voc_corr, method = "circle", type = "upper", tl.col = "black", tl.srt = 45, tl.cex = .3)
corrplot(voc_corr, method = "circle", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45, tl.cex = .3)
```


For females
```{r}
# Correlation
x <- voc_ordination_data %>% 
  filter(ssex == "f")
voc_corr <- cor(x[4:78])
par(mfrow = c(1,2))
corrplot(voc_corr, method = "circle", type = "upper", tl.col = "black", tl.srt = 45, tl.cex = .3)
corrplot(voc_corr, method = "circle", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45, tl.cex = .3)
```
